### 2024-03-17 paper/paper.tex

\documentclass[a4paper,12pt]{article}
\begin{document}

Hello World

\end{document}

### 2024-03-18 draw_orbits.py
add note about linux in the readme

@@ -2,7 +2,7 @@
 
 import physics_simulation as sim
 import matplotlib.pyplot as plt
-import matplotlib.animation as animation  
+import matplotlib.animation as animation
 import json
 import numpy as np
 
@@ -16,44 +16,45 @@ MAX_T = 31536000
 
 fig, ax = plt.subplots()
 
-ax.set_xlim([-258000000000, 258000000000])
-ax.set_ylim([-258000000000, 258000000000]) 
+# ax.set_xlim([-258000000000, 258000000000])
+# ax.set_ylim([-258000000000, 258000000000])
 
 scatter = ax.scatter([], [])
 
+
 # called for frame i
 def animate(i):
-    # t is a parameter which varies with the frame number 
+    # t is a parameter which varies with the frame number
     # 2 days per frame
-    t =  2 * SECONDS_IN_DAY * i  
+    # t = 2 * SECONDS_IN_DAY * i
 
     # slow down timer to see moon orbit earth
-    # t = 3600 * i
+    t = 3600 * i
 
     x_data = []
     y_data = []
-    
+
     # grab the data from physics_simulation.py
     data = sim.get_stats(t)
-    
-    for datum in data:       
-        # appending values to the previously  
-        # empty x and y data holders  
-        x_data.append(datum["x"])  
+
+    for datum in data:
+        # appending values to the previously
+        # empty x and y data holders
+        x_data.append(datum["x"])
         y_data.append(datum["y"])
 
     # uncomment to see moon orbit earth
-    # ax.set_xlim([data[1]["x"] - EARTH_ORBIT_RAD, data[1]["x"] + EARTH_ORBIT_RAD])
-    # ax.set_ylim([data[1]["y"] - EARTH_ORBIT_RAD, data[1]["y"] + EARTH_ORBIT_RAD]) 
+    ax.set_xlim([data[1]["x"] - EARTH_ORBIT_RAD, data[1]["x"] + EARTH_ORBIT_RAD])
+    ax.set_ylim([data[1]["y"] - EARTH_ORBIT_RAD, data[1]["y"] + EARTH_ORBIT_RAD])
 
     # add it to animation
-    scatter.set_offsets(np.column_stack((x_data, y_data)))  
-    return scatter,
+    scatter.set_offsets(np.column_stack((x_data, y_data)))
+    return (scatter,)
 
 
 # bootstrap for animated python file
 def animated():
[…]

### 2024-03-18 draw_orbits.py
revert change to draw-orbits

@@ -2,7 +2,7 @@
 
 import physics_simulation as sim
 import matplotlib.pyplot as plt
-import matplotlib.animation as animation
+import matplotlib.animation as animation  
 import json
 import numpy as np
 
@@ -16,45 +16,44 @@ MAX_T = 31536000
 
 fig, ax = plt.subplots()
 
-# ax.set_xlim([-258000000000, 258000000000])
-# ax.set_ylim([-258000000000, 258000000000])
+ax.set_xlim([-258000000000, 258000000000])
+ax.set_ylim([-258000000000, 258000000000]) 
 
 scatter = ax.scatter([], [])
 
-
 # called for frame i
 def animate(i):
-    # t is a parameter which varies with the frame number
+    # t is a parameter which varies with the frame number 
     # 2 days per frame
-    # t = 2 * SECONDS_IN_DAY * i
+    t =  2 * SECONDS_IN_DAY * i  
 
     # slow down timer to see moon orbit earth
-    t = 3600 * i
+    # t = 3600 * i
 
     x_data = []
     y_data = []
-
+    
     # grab the data from physics_simulation.py
     data = sim.get_stats(t)
-
-    for datum in data:
-        # appending values to the previously
-        # empty x and y data holders
-        x_data.append(datum["x"])
+    
+    for datum in data:       
+        # appending values to the previously  
+        # empty x and y data holders  
+        x_data.append(datum["x"])  
         y_data.append(datum["y"])
 
     # uncomment to see moon orbit earth
-    ax.set_xlim([data[1]["x"] - EARTH_ORBIT_RAD, data[1]["x"] + EARTH_ORBIT_RAD])
-    ax.set_ylim([data[1]["y"] - EARTH_ORBIT_RAD, data[1]["y"] + EARTH_ORBIT_RAD])
+    # ax.set_xlim([data[1]["x"] - EARTH_ORBIT_RAD, data[1]["x"] + EARTH_ORBIT_RAD])
+    # ax.set_ylim([data[1]["y"] - EARTH_ORBIT_RAD, data[1]["y"] + EARTH_ORBIT_RAD]) 
 
     # add it to animation
-    scatter.set_offsets(np.column_stack((x_data, y_data)))
-    return (scatter,)
+    scatter.set_offsets(np.column_stack((x_data, y_data)))  
+    return scatter,
 
 
 # bootstrap for animated python file
 def animated():
[…]

### 2024-03-20 physics_simulation.py
transmission_timing

@@ -1,6 +1,9 @@
 import math
 import json
 
+# Speed of light m/s
+C = 299792458.0
+
 # sun is at polar coordinates 0, 0
 
 # average sun radius:  695 508 000 m 
@@ -160,7 +163,9 @@ def get_stats(t: int):
             entity_stats = get_entity_stats(t, entity, stats)
             stats.append(entity_stats)
 
-    return get_connections(stats)
+    stats = get_connections(stats)
+
+    return stats
 
 
 # given two points to make a line, and a point
@@ -237,14 +242,35 @@ def get_connections(stats):
                         if dist < entity_blocking["radius"]:
                             blocking = True
 
+                if not blocking:
+                    dist = dist_between_points(entity_receiving["x"], entity_receiving["y"], entity_sending["x"], entity_sending["y"])
+                    trans_time = transmission_time(dist)
+                else:
+                    dist = None
+                    trans_time = None
+
                 entity_sending["connections"].append(
                     {
                         "id": entity_receiving["id"],
-                        "connected": not blocking
+                        "connected": not blocking,
+                        "distance": dist,
+                        "trans_time": trans_time,
                     }
                 )
     
     return stats
+
+
+# sending messages at speed of light
+# given a distance how long does it take
+def transmission_time(dist):
+    return dist / C
+
+
+# pythagorean theorem
+def dist_between_points(x1, y1, x2, y2):
+    return math.sqrt((x2 - x1) ** 2 + (y2 - y1) ** 2)
+
 
 # for testing
 # print(json.dumps(get_stats(0), indent=4))

### 2024-03-20 physics_simulation.py
added names

@@ -29,6 +29,7 @@ C = 299792458.0
 # randomize initial positions
 
 # id: the id for the entity
+# name: the english name of the entity
 # or: orbital radius of the entity (m)
 # orb_id: the id of the entity which this entity orbits
 #         The sun will be set to -1 (no index), since it is not orbiting any modeled entities
@@ -39,6 +40,7 @@ C = 299792458.0
 initial_pos = [
     { # Sun
         "id": 0,
+        "name": "Sun",
         "orbital_radius": 0,
         "orb_id": -1,
         "period": 1,
@@ -47,6 +49,7 @@ initial_pos = [
         "orbital_direction": 1,
     },{ # Earth
         "id": 1,
+        "name": "Earth",
         "orbital_radius": 149600000000,
         "orb_id": 0,
         "period": 31558149,
@@ -55,6 +58,7 @@ initial_pos = [
         "orbital_direction": 1,
     },{ # Mars
         "id": 2,
+        "name": "Mars",
         "orbital_radius": 228000000000,
         "orb_id": 0,
         "period": 59355072,
@@ -63,6 +67,7 @@ initial_pos = [
         "orbital_direction": 1,
     },{ # Moon
         "id": 3,
+        "name": "The Moon",
         "orbital_radius": 385000000,
         "orb_id": 1,
         "period": 361592,
@@ -71,6 +76,7 @@ initial_pos = [
         "orbital_direction": 1,
     },{ # ISS
         "id": 4,
+        "name": "ISS",
         "orbital_radius": 6371000 + 400000, # Earth radius + 400km
         "orb_id": 1,
         "period": 5580,
@@ -79,6 +85,7 @@ initial_pos = [
         "orbital_direction": 1,
     },{ # Mars Orbiter
         "id": 5,
+        "name": "Mars Orbiter",
         "orbital_radius": 3389500 + 400000, # Mars radius + 400km
         "orb_id": 2,
         "period": 7200,
@@ -127,6 +134,7 @@ def get_entity_stats(t, entity_stats, stat_list):
 
     return {
         "id": entity_stats["id"],
+        "name": entity_stats["name"],
         "orbital_radius": entity_stats["orbital_radius"],
         "orb_id": entity_stats["orb_id"],
         "period": entity_stats["period"],
@@ -149,6 +157,7 @@ def get_stats(t: int):
             stats.append(
                 {
                     "id": 0,
+                    "name": "Sun",
                     "orbital_radius": 0,
                     "orb_id": -1,
                     "period": 1,
[…]

### 2024-03-20 paper/paper.tex
physics engine latex

@@
 \documentclass[a4paper,12pt]{article}
+
+\title{Analyzing Network Transfer Protocols on a Simulated Interplanetary Internet}
+\author{Cody Auch, Mark Lysack, Jacob Janzen}
+\date{\today}
+
 \begin{document}
+\maketitle
 
-Hello World
+\section{Abstract}
+
+\section{Introduction}
+
+In the past decade interest in returning back to space has increased internationally. Many organizations 
+have started sending remote vehicles into space for the first time \cite{(indian space program, spaceX, ect)}
+and there are talks of a permanent lunar base being created. With an increased amount of satellites and potentially 
+permanent bases outside of Earths orbit, the demand on the interplanetary network (IPN) is increased. The IPN has a 
+series of unique challenges such as a high amount of error, constantly changing network topology, and long 
+propagation delay. The solution to these problems is implementing a delay tolerant network. While space organizations 
+have maintained the IPN with a set of standard protocols, these may not be sufficient for an expanding network.
+The simulator proposed shows how an IPN can handle different stages of development of an IPN as well as increased 
+throughput requirements.
+
+\section{Previous Work}
+
+\section{Methodology}
+
+\subsection{Physics Engine}
+
+A significant challenge in creating a IPN is the rapidly changing topology. To simulate 
+these rapid changing conditions a physics engine was created to model the orbits of several 
+entities in the solar system. Several assumptions were made in the simulator for the sake of 
+simplicity as the goal of the simulator is not to create a perfect model of the solar system 
+and the entities inside of the solar system, but rather conditions which a IPN would exist in.
+
+The first assumption made was that all orbits are circular. Realistically, orbits are oval 
+shaped, with the two focal points. A result of this assumption is the elimination of any 
+difference in orbital speed due to Kepler's third law. The orbital radius was chosen by taking
+the average orbital radius of the real life celestial bodies.
+For example Earth's average orbital
+radius is 149.6 billion meters \cite{}, and therefore orbits around the sun in a circular orbit
+of radius 149.6 billion meters.  Another result of this assumption 
+is that orbits are considered to be centered around the body which it orbits. The Mars orbit 
+has a focal point that is further in space, making it more oval shaped than the earth orbit. \cite{} 
+While this could make changes to a 
+simulated network in some very specific situations, the primary behaviour of a solar system 
+is maintained, like retrograde orbits, periodical alignment, predictable behaviour, and 
+recursive orbits of sub orbiting object.
+
+All entities are assumed to have a circular shape, with a constant radius. Due to the scale of the 
+objects, both large and small, this has a very minor effect.
+
+The whole physics simulation has been done on a 2D plane. While planetary orbits were not perfectly
+aligned, on the astronomical scale which we are working on, assuming that all their orbits are 
+contained on the same plane will not effect the simulation significantly. Where this could be an issue 
+is for when there are more than two satellites orbiting the same entity, with a relatively low orbit. For
+example satellites orbiting Earth. While there are some complex satellite interactions which are removed 
+it does still simulate satellites having to form a network around the entity it is 
+orbiting, it just requires less satellites then if it needed to form a 3D network.
+
+Messages between entities were assumed to be sent at the speed of light, as all forms of radio, lasers,
+or other wireless messages all are sent at the speed of light. Since propagation delay is much higher 
+then terrestrial networks, other delays (TODO specify which delays) were ignored as being negligible.
+
+Based on these simplifications, a physical simulation with several goals. To calculate if a line of 
+sight could be made between two communicating entities, that is if a message can be sent, to calculate 
+the error rate when sending a message, and to calculate the message propagation delay for the message to 
+be sent. Line of sight was achieved using simple geometry calculations. (See appendix) All entities are able to block 
+signals if the ray passes through the radius of the entity. (TODO Cody inserts bit about error calculation)
+Propagation delay is a simple function of distance and the speed of light. The input of the physics engine 
+is a time $t$ from some preset initial conditions which outputs the state of all entities.
 

### 2024-03-21 physics_simulation.py
Added FSPL calculation

@@ -4,6 +4,9 @@ import json
 # Speed of light m/s
 C = 299792458.0
 
+# Transmission Frequency (Hz)
+T_FREQ = (30*10)**9
+
 # sun is at polar coordinates 0, 0
 
 # average sun radius:  695 508 000 m 
@@ -281,6 +284,21 @@ def transmission_time(dist):
 # pythagorean theorem
 def dist_between_points(x1, y1, x2, y2):
     return math.sqrt((x2 - x1) ** 2 + (y2 - y1) ** 2)
+
+
+# Calculate the free space path loss between 2 points
+#
+def free_space_path_loss(x1, y1, x2, y2) -> float:
+    # Define the transmission frequency (30GHz, middle of Ka-band) and calculate wavelength
+    wavelength = C / T_FREQ
+
+    # Distance between the points 
+    distance = dist_between_points(x1, y1, x2, y2) / 1000 
+
+    # Assuming transmission/reception via isotropic antennas
+    loss_ratio = ((4 * math.pi * distance)/wavelength) ** 2
+    
+    return loss_ratio
 
 
 # for testing

### 2024-03-28 network_sim.py
take down old nodes

@@ -16,23 +16,46 @@ NodeContainer create_node_pair(NodeContainer nodes, int i, int j) {
     return NodeContainer(nodes.Get(i), nodes.Get(j));
 }
 
-void node_set_down(NodeContainer c, int index, int interface) {
+void set_down(NodeContainer c, int index, int interface) {
     Ptr<Node> n = c.Get(index);
     Ptr<Ipv4> ipv4 = n->GetObject<Ipv4>();
     ipv4->SetDown(interface);
 }
 
-void node_set_up(NodeContainer c, int index, int interface) {
+void set_up(NodeContainer c, int index, int interface) {
     Ptr<Node> n = c.Get(index);
     Ptr<Ipv4> ipv4 = n->GetObject<Ipv4>();
     ipv4->SetUp(interface);
+}
+
+void print_routing_table() {
+    Ipv4GlobalRoutingHelper globalRouting;
+    Ptr<OutputStreamWrapper> routingStream = Create<OutputStreamWrapper> (&std::cout);
+    globalRouting.PrintRoutingTableAllAt (Seconds(0.1), routingStream );
+}
+
+int get_num_devices(NodeContainer c, int index) {
+    Ptr<Node> n = c.Get(index);
+
+    return n->GetNDevices();
+}
+
+int get_netdevice_node_index(NodeContainer c, int node_index, int device_index) {
+    Ptr<Channel> ch = c.Get(node_index)->GetDevice(device_index)->GetChannel();
+    int i1 = ch->GetDevice(0)->GetNode()->GetId();
+    int i2 = ch->GetDevice(1)->GetNode()->GetId();
+
+    return node_index == i1 ? i2 : i1;
 }
 """
 )
 
 create_node_pair = ns.cppyy.gbl.create_node_pair
-set_down = ns.cppyy.gbl.node_set_down
-set_up = ns.cppyy.gbl.node_set_up
+set_down = ns.cppyy.gbl.set_down
+set_up = ns.cppyy.gbl.set_up
+print_routing_table = ns.cppyy.gbl.print_routing_table
+get_num_devices = ns.cppyy.gbl.get_num_devices
+get_netdevice_node_index = ns.cppyy.gbl.get_netdevice_node_index
 
 
 @dataclass
@@ -41,10 +64,14 @@ class Topology:
     channel_table: "List[List[NetDeviceContainer|None]]"
     ip_table: "List[List[Ipv4InterfaceContainer|None]]"
 
+@dataclass
+class ConnectionData:
[…]

### 2024-03-28 network_sim.py
add channel delay

@@ -47,6 +47,12 @@ int get_netdevice_node_index(NodeContainer c, int node_index, int device_index) 
 
     return node_index == i1 ? i2 : i1;
 }
+
+void set_channel_delay(NodeContainer c, int node_index, int device_index, double seconds) {
+    Ptr<Channel> ch = c.Get(node_index)->GetDevice(device_index)->GetChannel();
+
+    ch->SetAttribute("Delay", TimeValue(Seconds(seconds)));
+}
 """
 )
 
@@ -56,7 +62,7 @@ set_up = ns.cppyy.gbl.set_up
 print_routing_table = ns.cppyy.gbl.print_routing_table
 get_num_devices = ns.cppyy.gbl.get_num_devices
 get_netdevice_node_index = ns.cppyy.gbl.get_netdevice_node_index
-
+set_channel_delay = ns.cppyy.gbl.set_channel_delay
 
 @dataclass
 class Topology:
@@ -155,10 +161,11 @@ def update_topology(topology: Topology, time: int) -> None:
                 conn_id = get_netdevice_node_index(topology.nodes, id, i)
                 if conn_id in conns and conns[conn_id].connected:
                     set_up(topology.nodes, id, i)
+                    set_channel_delay(topology.nodes, id, i, conns[conn_id].trans_time)
                 else:
                     set_down(topology.nodes, id, i)
 
-        # TODO update delay, error rate, etc
+        # TODO update error rate
 
     ns.internet.Ipv4GlobalRoutingHelper.RecomputeRoutingTables()
 
@@ -174,7 +181,7 @@ def simulate():
     ns.core.LogComponentEnable("PacketSink", ns.core.LOG_LEVEL_INFO)
 
     topology = create_topology()
-    install_onoff_app(topology, 1, 1, 5)
+    install_onoff_app(topology, 2, 1, 5)
     install_sink(topology, 5)
     update_topology(topology, 10000)
 

### 2024-03-28 network_sim.py
schedule state recomputation

@@ -10,6 +10,8 @@ Ipv4InterfaceContainer = ns.cppyy.gbl.ns3.Ipv4InterfaceContainer
 
 ns.cppyy.cppdef(
     """
+#include "CPyCppyy/API.h"
+
 using namespace ns3;
 
 NodeContainer create_node_pair(NodeContainer nodes, int i, int j) {
@@ -53,6 +55,11 @@ void set_channel_delay(NodeContainer c, int node_index, int device_index, double
 
     ch->SetAttribute("Delay", TimeValue(Seconds(seconds)));
 }
+
+// Schedule seems to require CPP functions. This gets around that by calling the python function from CPP
+void cpp_update_topology() {
+    CPyCppyy::Eval("update_topology()");
+}
 """
 )
 
@@ -63,6 +70,12 @@ print_routing_table = ns.cppyy.gbl.print_routing_table
 get_num_devices = ns.cppyy.gbl.get_num_devices
 get_netdevice_node_index = ns.cppyy.gbl.get_netdevice_node_index
 set_channel_delay = ns.cppyy.gbl.set_channel_delay
+cpp_update_topology = ns.cppyy.gbl.cpp_update_topology
+
+global_time = 0
+global_topology = None
+
+TIME_STEP = 60
 
 @dataclass
 class Topology:
@@ -141,15 +154,19 @@ def install_sink(topology: Topology, index: int) -> None:
     apps.Start(ns.core.Seconds(1.0))
     apps.Stop(ns.core.Seconds(10.0))
 
-def update_topology(topology: Topology, time: int) -> None:
-    entities = get_stats(time)
-    print(json.dumps(entities, indent=2))
+def update_topology() -> None:
+    if global_topology is None:
+        return
+
+    global global_time
+
+    entities = get_stats(global_time)
     for entity in entities:
         id = entity["id"]
         if not entity["can_connect"]:
             # take down each interface starting from index 1 (index 0 is loopback)
-            for i in range(1, get_num_devices(topology.nodes, id)):
-                set_down(topology.nodes, id, i)
+            for i in range(1, get_num_devices(global_topology.nodes, id)):
+                set_down(global_topology.nodes, id, i)
         else:
             conns = {}
             for i, conn in enumerate(entity["connections"]):
@@ -157,20 +174,24 @@ def update_topology(topology: Topology, time: int) -> None:
                     connected=conn["connected"],
                     trans_time=conn["trans_time"],
                 )
[…]

### 2024-03-28 network_sim.py
run for one day and slow down data rate

@@ -77,6 +77,8 @@ global_topology = None
 
 TIME_STEP = 60
 
+SIM_LENGTH = 60 * 60 * 24
+
 @dataclass
 class Topology:
     nodes: "NodeContainer"
@@ -140,19 +142,22 @@ def install_onoff_app(topology: Topology, index: int, ch_i: int, ch_j: int) -> N
         return
 
     onoff = ns.applications.OnOffHelper("ns3::UdpSocketFactory", ns.network.InetSocketAddress(address.GetAddress(1), port).ConvertTo())
-    onoff.SetConstantRate(ns.network.DataRate("2kbps"))
-    onoff.SetAttribute("PacketSize", ns.core.UintegerValue(50))
+
+    # one packet per TIME_STEP
+    rate = 1 / TIME_STEP
+    onoff.SetConstantRate(ns.network.DataRate(f"{rate}kbps"))
+    onoff.SetAttribute("PacketSize", ns.core.UintegerValue(1024))
 
     apps = onoff.Install(topology.nodes.Get(index))
-    apps.Start(ns.core.Seconds(1.0))
-    apps.Stop(ns.core.Seconds(10.0))
+    apps.Start(ns.core.Seconds(0.0))
+    apps.Stop(ns.core.Seconds(SIM_LENGTH))
 
 def install_sink(topology: Topology, index: int) -> None:
     port = 9 # Discard port (RFC 863)
     sink = ns.applications.PacketSinkHelper("ns3::UdpSocketFactory", ns.InetSocketAddress(ns.Ipv4Address.GetAny(), port).ConvertTo())
     apps = sink.Install(topology.nodes.Get(index))
-    apps.Start(ns.core.Seconds(1.0))
-    apps.Stop(ns.core.Seconds(10.0))
+    apps.Start(ns.core.Seconds(0.0))
+    apps.Stop(ns.core.Seconds(SIM_LENGTH))
 
 def update_topology() -> None:
     if global_topology is None:
@@ -207,7 +212,6 @@ def simulate():
     install_sink(topology, 5)
     cpp_update_topology()
 
-    # TODO schedule recomputation of state
     print_routing_table()
 
 

### 2024-03-28 network_sim.py
reformat network_sim

@@ -79,11 +79,13 @@ TIME_STEP = 60
 
 SIM_LENGTH = 60 * 60 * 24
 
+
 @dataclass
 class Topology:
     nodes: "NodeContainer"
     channel_table: "List[List[NetDeviceContainer|None]]"
     ip_table: "List[List[Ipv4InterfaceContainer|None]]"
+
 
 @dataclass
 class ConnectionData:
@@ -135,13 +137,17 @@ def create_topology() -> Topology:
         ip_table=ip_table,
     )
 
+
 def install_onoff_app(topology: Topology, index: int, ch_i: int, ch_j: int) -> None:
-    port = 9 # Discard port (RFC 863)
+    port = 9  # Discard port (RFC 863)
     address = topology.ip_table[ch_i][ch_j]
     if address is None:
         return
 
-    onoff = ns.applications.OnOffHelper("ns3::UdpSocketFactory", ns.network.InetSocketAddress(address.GetAddress(1), port).ConvertTo())
+    onoff = ns.applications.OnOffHelper(
+        "ns3::UdpSocketFactory",
+        ns.network.InetSocketAddress(address.GetAddress(1), port).ConvertTo(),
+    )
 
     # one packet per TIME_STEP
     rate = 1 / TIME_STEP
@@ -152,12 +158,17 @@ def install_onoff_app(topology: Topology, index: int, ch_i: int, ch_j: int) -> N
     apps.Start(ns.core.Seconds(0.0))
     apps.Stop(ns.core.Seconds(SIM_LENGTH))
 
+
 def install_sink(topology: Topology, index: int) -> None:
-    port = 9 # Discard port (RFC 863)
-    sink = ns.applications.PacketSinkHelper("ns3::UdpSocketFactory", ns.InetSocketAddress(ns.Ipv4Address.GetAny(), port).ConvertTo())
+    port = 9  # Discard port (RFC 863)
+    sink = ns.applications.PacketSinkHelper(
+        "ns3::UdpSocketFactory",
+        ns.InetSocketAddress(ns.Ipv4Address.GetAny(), port).ConvertTo(),
+    )
     apps = sink.Install(topology.nodes.Get(index))
     apps.Start(ns.core.Seconds(0.0))
     apps.Stop(ns.core.Seconds(SIM_LENGTH))
+
 
 def update_topology() -> None:
     if global_topology is None:
@@ -183,7 +194,9 @@ def update_topology() -> None:
[…]

### 2024-03-28 network_sim.py
don't print routing table

@@ -225,8 +225,6 @@ def simulate():
     install_sink(topology, 5)
     cpp_update_topology()
 
-    print_routing_table()
-
     ns.core.Simulator.Schedule(ns.core.Seconds(60.0), cpp_update_topology)
 
     ns.core.Simulator.Run()

### 2024-03-28 network_sim.py
remove todo and start at more interesting time

@@ -72,7 +72,7 @@ get_netdevice_node_index = ns.cppyy.gbl.get_netdevice_node_index
 set_channel_delay = ns.cppyy.gbl.set_channel_delay
 cpp_update_topology = ns.cppyy.gbl.cpp_update_topology
 
-global_time = 0
+global_time = 10000
 global_topology = None
 
 TIME_STEP = 60
@@ -116,7 +116,6 @@ def create_topology() -> Topology:
             else:
                 np = create_node_pair(nodes, i, j)
 
-                # TODO set attributes based on real statistics
                 p2p = ns.point_to_point.PointToPointHelper()
                 p2p.SetDeviceAttribute("DataRate", ns.core.StringValue("5Mbps"))
                 ch = p2p.Install(np)
@@ -221,7 +220,7 @@ def simulate():
 
     topology = create_topology()
     global_topology = topology
-    install_onoff_app(topology, 2, 1, 5)
+    install_onoff_app(topology, 1, 1, 5)
     install_sink(topology, 5)
     cpp_update_topology()
 

### 2024-03-28 network_sim.py
fix rate

@@ -94,7 +94,8 @@ class ConnectionData:
 
 
 def create_topology() -> Topology:
-    entities = get_stats(0)
+    entities = get_stats(10000)
+    print(json.dumps(entities, indent=2))
     nodes = ns.network.NodeContainer()
     nodes.Create(len(entities))
 
@@ -149,7 +150,7 @@ def install_onoff_app(topology: Topology, index: int, ch_i: int, ch_j: int) -> N
     )
 
     # one packet per TIME_STEP
-    rate = 1 / TIME_STEP
+    rate = 8 / TIME_STEP
     onoff.SetConstantRate(ns.network.DataRate(f"{rate}kbps"))
     onoff.SetAttribute("PacketSize", ns.core.UintegerValue(1024))
 

### 2024-03-28 network_sim.py
fix (most of) the linter issues

@@ -1,7 +1,12 @@
-from typing import List
+"""network_sim
+
+run a simulation of interplanetary network
+"""
+
 import json
-from ns import ns
 from dataclasses import dataclass
+from typing import List
+from ns import ns
 from physics_simulation import get_stats
 
 NodeContainer = ns.cppyy.gbl.ns3.NodeContainer
@@ -56,7 +61,8 @@ void set_channel_delay(NodeContainer c, int node_index, int device_index, double
     ch->SetAttribute("Delay", TimeValue(Seconds(seconds)));
 }
 
-// Schedule seems to require CPP functions. This gets around that by calling the python function from CPP
+// Schedule seems to require CPP functions.
+// This gets around that by calling the python function from CPP
 void cpp_update_topology() {
     CPyCppyy::Eval("update_topology()");
 }
@@ -72,8 +78,8 @@ get_netdevice_node_index = ns.cppyy.gbl.get_netdevice_node_index
 set_channel_delay = ns.cppyy.gbl.set_channel_delay
 cpp_update_topology = ns.cppyy.gbl.cpp_update_topology
 
-global_time = 10000
-global_topology = None
+GLOBAL_TIME = 10000
+GLOBAL_TOPOLOGY = None
 
 TIME_STEP = 60
 
@@ -82,6 +88,12 @@ SIM_LENGTH = 60 * 60 * 24
 
 @dataclass
 class Topology:
+    """Topology
+
+    object containing data about Topology
+    (we don't really need this tbh)
+    """
+
     nodes: "NodeContainer"
     channel_table: "List[List[NetDeviceContainer|None]]"
     ip_table: "List[List[Ipv4InterfaceContainer|None]]"
@@ -89,11 +101,19 @@ class Topology:
 
 @dataclass
 class ConnectionData:
+    """ConnectionData
+
+    object containing data about a connection"""
+
     connected: bool
     trans_time: float
 
 
 def create_topology() -> Topology:
+    """create_topology
+
+    create a network topology from entities
+    """
     entities = get_stats(10000)
     print(json.dumps(entities, indent=2))
     nodes = ns.network.NodeContainer()
@@ -139,6 +159,15 @@ def create_topology() -> Topology:
 
 
 def install_onoff_app(topology: Topology, index: int, ch_i: int, ch_j: int) -> None:
+    """install_onoff_app
+
+    install an onoff application on a node
+
+    topology - the topology to install on
+    index    - the node to install on
+    ch_i     - index i to send to
+    ch_j     - index j to send to
+    """
[…]

### 2024-03-28 network_sim.py
actually comment the code

@@ -19,34 +19,40 @@ ns.cppyy.cppdef(
 
 using namespace ns3;
 
+// create a NodeContainer with two nodes
 NodeContainer create_node_pair(NodeContainer nodes, int i, int j) {
     return NodeContainer(nodes.Get(i), nodes.Get(j));
 }
 
+// set an interface to be down
 void set_down(NodeContainer c, int index, int interface) {
     Ptr<Node> n = c.Get(index);
     Ptr<Ipv4> ipv4 = n->GetObject<Ipv4>();
     ipv4->SetDown(interface);
 }
 
+// set an interface to be up
 void set_up(NodeContainer c, int index, int interface) {
     Ptr<Node> n = c.Get(index);
     Ptr<Ipv4> ipv4 = n->GetObject<Ipv4>();
     ipv4->SetUp(interface);
 }
 
+// print the global routing table
 void print_routing_table() {
     Ipv4GlobalRoutingHelper globalRouting;
     Ptr<OutputStreamWrapper> routingStream = Create<OutputStreamWrapper> (&std::cout);
     globalRouting.PrintRoutingTableAllAt (Seconds(0.1), routingStream );
 }
 
+// get the number of devices
 int get_num_devices(NodeContainer c, int index) {
     Ptr<Node> n = c.Get(index);
 
     return n->GetNDevices();
 }
 
+// get the index of a netdevice
 int get_netdevice_node_index(NodeContainer c, int node_index, int device_index) {
     Ptr<Channel> ch = c.Get(node_index)->GetDevice(device_index)->GetChannel();
     int i1 = ch->GetDevice(0)->GetNode()->GetId();
@@ -55,13 +61,14 @@ int get_netdevice_node_index(NodeContainer c, int node_index, int device_index) 
     return node_index == i1 ? i2 : i1;
 }
 
+// set the channel delay on a channel
 void set_channel_delay(NodeContainer c, int node_index, int device_index, double seconds) {
     Ptr<Channel> ch = c.Get(node_index)->GetDevice(device_index)->GetChannel();
 
     ch->SetAttribute("Delay", TimeValue(Seconds(seconds)));
 }
 
-// Schedule seems to require CPP functions.
+// Schedule seems to require CPP functions
 // This gets around that by calling the python function from CPP
 void cpp_update_topology() {
     CPyCppyy::Eval("update_topology()");
@@ -69,6 +76,7 @@ void cpp_update_topology() {
 """
 )
 
+# shorten names of cpp functions
[…]

### 2024-03-29 network_sim.py
add tracing

@@ -124,6 +124,10 @@ def create_topology() -> Topology:
 
     create a network topology from entities
     """
+
+    ascii = ns.network.AsciiTraceHelper()
+    stream = ascii.CreateFileStream("network-sim.tr")
+
     # get the number of nodes in the network
     entities = get_stats(0)
     nodes = ns.network.NodeContainer()
@@ -152,6 +156,7 @@ def create_topology() -> Topology:
 
                 p2p = ns.point_to_point.PointToPointHelper()
                 p2p.SetDeviceAttribute("DataRate", ns.core.StringValue("5Mbps"))
+                p2p.EnableAsciiAll(stream)
                 ch = p2p.Install(np)
                 channel_row.append(ch)
 

### 2024-03-29 network_sim.py
add interface mapping table

@@ -4,6 +4,7 @@ run a simulation of interplanetary network
 """
 
 import json
+import csv
 from dataclasses import dataclass
 from typing import List
 from ns import ns
@@ -104,6 +105,7 @@ class Topology:
     """
 
     nodes: "NodeContainer"
+    num_nodes: int
     channel_table: "List[List[NetDeviceContainer|None]]"
     ip_table: "List[List[Ipv4InterfaceContainer|None]]"
 
@@ -131,7 +133,8 @@ def create_topology() -> Topology:
     # get the number of nodes in the network
     entities = get_stats(0)
     nodes = ns.network.NodeContainer()
-    nodes.Create(len(entities))
+    num_nodes = len(entities)
+    nodes.Create(num_nodes)
 
     # install internet stack on each node
     internet = ns.internet.InternetStackHelper()
@@ -171,6 +174,7 @@ def create_topology() -> Topology:
 
     return Topology(
         nodes=nodes,
+        num_nodes=num_nodes,
         channel_table=channel_table,
         ip_table=ip_table,
     )
@@ -265,6 +269,15 @@ def update_topology() -> None:
     GLOBAL_TIME += TIME_STEP
 
 
+def map_interfaces(topology: Topology):
+    interface_map = []
+    for i in range(topology.num_nodes):
+        for j in range(1, get_num_devices(topology.nodes, i)):
+            interface_map.append([i, j, get_netdevice_node_index(topology.nodes, i, j)])
+
+    return interface_map
+
+
 def simulate() -> None:
     """simulate
 
@@ -290,6 +303,14 @@ def simulate() -> None:
 
     # schedule recomputation of topology once a minute
     ns.core.Simulator.Schedule(ns.core.Seconds(60.0), cpp_update_topology)
+
+    print_routing_table()
+
+    interface_map = map_interfaces(topology)
+    with open("interface_mapping.csv", "w+") as im_csv:
+        csv_writer = csv.writer(im_csv, delimiter=",")
+        csv_writer.writerow(["Input Node", "Interface", "Output Node"])
+        csv_writer.writerows(interface_map)
 
     # run the simulator
     ns.core.Simulator.Run()

### 2024-03-30 physics_simulation.py
Added error rate to stats object

@@ -258,9 +258,11 @@ def get_connections(stats):
                 if not blocking:
                     dist = dist_between_points(entity_receiving["x"], entity_receiving["y"], entity_sending["x"], entity_sending["y"])
                     trans_time = transmission_time(dist)
+                    err_rate = get_error_rate(entity_sending, entity_receiving)
                 else:
                     dist = None
                     trans_time = None
+                    err_rate = None
 
                 entity_sending["connections"].append(
                     {
@@ -269,6 +271,7 @@ def get_connections(stats):
                         "connected": not blocking,
                         "distance": dist,
                         "trans_time": trans_time,
+                        "error_rate": err_rate
                     }
                 )
     
@@ -287,19 +290,23 @@ def dist_between_points(x1, y1, x2, y2):
 
 
 # Calculate the free space path loss between 2 points
-#
+# https://en.wikipedia.org/wiki/Free-space_path_loss#Free-space_path_loss_formula
 def free_space_path_loss(x1, y1, x2, y2) -> float:
     # Define the transmission frequency (30GHz, middle of Ka-band) and calculate wavelength
     wavelength = C / T_FREQ
 
-    # Distance between the points 
-    distance = dist_between_points(x1, y1, x2, y2) / 1000 
+    # Distance between the points in m
+    distance = dist_between_points(x1, y1, x2, y2)
 
     # Assuming transmission/reception via isotropic antennas
-    loss_ratio = ((4 * math.pi * distance)/wavelength) ** 2
+    loss_ratio = (wavelength/(4 * math.pi * distance)) ** 2
     
     return loss_ratio
 
+# Get the rate of transmission error between two satelites
+def get_error_rate(sender, receiver) -> float:
+    assert("x" in sender.keys() and "y" in sender.keys() and "x" in receiver.keys() and "y" in receiver.keys())
+    return free_space_path_loss(sender["x"], sender["y"], receiver["x"], receiver["y"])
 
 # for testing
 # print(json.dumps(get_stats(0), indent=4))

### 2024-03-30 process_data.py
process data

@@ -1,0 +1,72 @@
+import re
+import pandas as pd
+
+
+def process_trace():
+    file1 = open('network-sim.tr', 'r')
+    Lines = file1.readlines()
+
+    processed_data = []
+
+    for line in Lines:
+        try:
+            TYPE_INDEX = 1
+            TIME_INDEX = 2
+            NODE_INDEX = 3
+            INTERFACE_INDEX = 4
+
+            res = re.compile("^([r\\+-]) ([0-9]*\.[0-9]*|[0-9]*[0-9]*) \/NodeList\/([0-9]*)\/DeviceList\/([0-9]*)").search(line)
+
+            type = res.group(TYPE_INDEX)
+            time = res.group(TIME_INDEX)
+            node = res.group(NODE_INDEX)
+            interface = res.group(INTERFACE_INDEX)
+
+            id = re.compile("(?: id )(.*?(?= ))").search(line).group(1)
+
+            processed_data.append([type, time, node, interface, id])
+
+        except Exception as e:
+            print("exception encountered")
+            print(line)
+            print(e)
+
+    
+    df = pd.DataFrame(processed_data, columns=["type", "time", "node1", "interface", "id"])
+    return df
+
+
+def create_stats_df():
+    trace_df = process_trace()
+    interface_mapping_df = pd.read_csv("interface_mapping.csv")
+
+    trace_df["node1"] = pd.to_numeric(trace_df["node1"])
+    trace_df["interface"] = pd.to_numeric(trace_df["interface"])
+
+    df = trace_df.merge(interface_mapping_df, left_on=["node1", "interface"], right_on=["Input Node", "Interface"], how="left")
+
+    df["node2"] = df["Output Node"]
+    
+    df = df[["node1", "interface", "node2", "type", "time", "id"]]
+
+    df = df.drop_duplicates()
+
+    # need to merge the df to itself, need to join them together
+    # merge on id?
+
+    df_queue = df[df["type"] == "+"]
+    df_dequeue = df[df["type"] == "-"]
+    df_type = df[df["type"] == "r"]
+
+    df = df_queue.merge(df_dequeue, how="left", on="id")
+    df = df.merge(df_type, how="left", on="id")
+
+    df[["node_sending", "node_receiving", "interface", "time_queued", "time_dequeued", "time_received", "id"]] = df[["node1", "node2", "interface", "time", "time_x", "time_y", "id"]]
+
[…]

### 2024-03-30 process_data.py
removed temp comment

@@ -1,7 +1,7 @@
 import re
 import pandas as pd
 
-
+# take trace file and turn it into a dataframe
 def process_trace():
     file1 = open('network-sim.tr', 'r')
     Lines = file1.readlines()
@@ -36,6 +36,8 @@ def process_trace():
     return df
 
 
+# based off of the trace file create a dataframe where each row is for 1 message id
+# has information on the time queued, time dequeued, time received, sending and receiving nodes
 def create_stats_df():
     trace_df = process_trace()
     interface_mapping_df = pd.read_csv("interface_mapping.csv")
@@ -52,7 +54,6 @@ def create_stats_df():
     df = df.drop_duplicates()
 
     # need to merge the df to itself, need to join them together
-    # merge on id?
 
     df_queue = df[df["type"] == "+"]
     df_dequeue = df[df["type"] == "-"]
@@ -69,4 +70,4 @@ def create_stats_df():
 
 
 # for testing
-print(create_stats_df())
+# print(create_stats_df())

### 2024-03-30 process_data.py
corrected column checks

@@ -1,6 +1,7 @@
 import re
 import pandas as pd
 
+
 # take trace file and turn it into a dataframe
 def process_trace():
     file1 = open('network-sim.tr', 'r')
@@ -62,7 +63,7 @@ def create_stats_df():
     df = df_queue.merge(df_dequeue, how="left", on="id")
     df = df.merge(df_type, how="left", on="id")
 
-    df[["node_sending", "node_receiving", "interface", "time_queued", "time_dequeued", "time_received", "id"]] = df[["node1", "node2", "interface", "time", "time_x", "time_y", "id"]]
+    df[["node_sending", "node_receiving", "interface", "time_queued", "time_dequeued", "time_received", "id"]] = df[["node1_x", "node2_x", "interface_x", "time_x", "time_y", "time", "id"]]
 
     df = df[["node_sending", "node_receiving", "interface", "time_queued", "time_dequeued", "time_received", "id"]]
 
@@ -70,4 +71,4 @@ def create_stats_df():
 
 
 # for testing
-# print(create_stats_df())
+print(create_stats_df())

### 2024-03-31 paper/paper.tex
Added a description of error rate calculation

@@
 
-\section{Abstract}
+\begin{abstract}
+    TODO insert abstract here
+\end{abstract}
 
@@
 be sent. Line of sight was achieved using simple geometry calculations. (See appendix) All entities are able to block 
-signals if the ray passes through the radius of the entity. (TODO Cody inserts bit about error calculation)
+signals if the ray passes through the radius of the entity.
+One of the largest factors contributing to transmission error is the distance  travelled by the electromagnetic signal between transmitting and receiving nodes.
+By generalizing Friis' transmission formula for isotropic emitters and receivers, we can obtain the ratio of signal intensity received to signal intensity transmitted, known as Free Space Path Loss (FSPL).
+We can use FSPL as our transmission error rate, giving the following expression: $Error Rate = (\frac{4 \pi d}{\lambda})^2$ where $d$ is the distance between nodes in metres and $\lambda$ is the wavelength of the electromagnetic wave used in transmission.
+As much of space communications utilizes the Ka-band of frequencies (27-40 GHz) we opted to use a transmission frequency of 30Ghz.
+The transmission wavelength can be then calculated according to $\lambda = \frac{c}{f}$, where $c$ is the speed of light and $f$ is the transmission frequency, giving a transmission wavelength of ~9.99mm.
 Propagation delay is a simple function of distance and the speed of light. The input of the physics engine 

### 2024-04-02 network_sim.py
add error rate to network simulation

@@ -69,6 +69,16 @@ void set_channel_delay(NodeContainer c, int node_index, int device_index, double
     ch->SetAttribute("Delay", TimeValue(Seconds(seconds)));
 }
 
+// set the error rate on a channel
+void set_channel_error(NodeContainer c, int node_index, int device_index, double rate) {
+    Ptr<NetDevice> ch = c.Get(node_index)->GetDevice(device_index);
+    Ptr<RateErrorModel> em = CreateObject<RateErrorModel>();
+    em->SetAttribute("ErrorRate", DoubleValue(rate));
+    em->SetUnit(RateErrorModel::ERROR_UNIT_BIT);
+
+    ch->SetAttribute("ReceiveErrorModel", PointerValue(em));
+}
+
 // Schedule seems to require CPP functions
 // This gets around that by calling the python function from CPP
 void cpp_update_topology() {
@@ -85,6 +95,7 @@ print_routing_table = ns.cppyy.gbl.print_routing_table
 get_num_devices = ns.cppyy.gbl.get_num_devices
 get_netdevice_node_index = ns.cppyy.gbl.get_netdevice_node_index
 set_channel_delay = ns.cppyy.gbl.set_channel_delay
+set_channel_error = ns.cppyy.gbl.set_channel_error
 cpp_update_topology = ns.cppyy.gbl.cpp_update_topology
 
 # global vars (we need global state to allow calling Python function from CPP)
@@ -119,6 +130,7 @@ class ConnectionData:
 
     connected: bool
     trans_time: float
+    error_rate: float
 
 
 def create_topology() -> Topology:
@@ -253,6 +265,7 @@ def update_topology() -> None:
                 conns[conn["id"]] = ConnectionData(
                     connected=conn["connected"],
                     trans_time=conn["trans_time"],
+                    error_rate=conn["error_rate"],
                 )
             for i in range(1, get_num_devices(GLOBAL_TOPOLOGY.nodes, id)):
                 conn_id = get_netdevice_node_index(GLOBAL_TOPOLOGY.nodes, id, i)
@@ -261,6 +274,7 @@ def update_topology() -> None:
                     set_channel_delay(
                         GLOBAL_TOPOLOGY.nodes, id, i, conns[conn_id].trans_time
                     )
+                    set_channel_error(GLOBAL_TOPOLOGY.nodes, id, i, conns[conn_id].error_rate)
                 else:
[…]

### 2024-04-02 network_sim.py
reformat file

@@ -274,7 +274,9 @@ def update_topology() -> None:
                     set_channel_delay(
                         GLOBAL_TOPOLOGY.nodes, id, i, conns[conn_id].trans_time
                     )
-                    set_channel_error(GLOBAL_TOPOLOGY.nodes, id, i, conns[conn_id].error_rate)
+                    set_channel_error(
+                        GLOBAL_TOPOLOGY.nodes, id, i, conns[conn_id].error_rate
+                    )
                 else:
                     set_down(GLOBAL_TOPOLOGY.nodes, id, i)
 

### 2024-04-03 paper/paper.tex
previous work and intro to interplanetary network

--- 2024-03-20 paper/paper.tex
@@
 The simulator proposed shows how an IPN can handle different stages of development of an IPN as well as increased 
-throughput requirements.
+throughput requirements. Before these technological changes can be implemented simulations need to be created to 
+lower the risk of failure given the high cost of failure. Often satellites will need to be in operation for years 
+or even decades. The hardware and software in satellites can often not be updated very well. Thus it will be very 
+relevant to investigate these issues well before they are needed.
 
 \section{Previous Work}
+
+In their 2017 Masters Thesis, Md Monjurul Islam Khan \cite{Khan2017} from University of Manitoba wrote on how 
+to create a satellite network
+simulation based on ns3. They used a point to point ns3 module to create a network of 2 levels of orbiting 
+satellites and fixed points on the earth which acts as uplinks to the network. The paper analyzed TCP 
+HighSpeed against TCP Vegas slow start. However this paper stops at expanding the network beyond earth orbit, 
+and rather focuses on Earth orbit communication.
+
+SNS3 is a simulation library which uses NS3 as a basis to create a framework which to build simulations for 
+geostationary satellites in the european continent. It does not expand to an interplanetary network.
+\cite{10.4108/icst.simutools.2014.254631}
+
+For non simulation work done on the interplanetary internet, the leading authority is Consultative Committee for 
+Space Data Systems (CCSDS). \cite{CCSDS.org} They maintain and publish the standards for interplanetary internet, including 
+the SCPS TP, a reliable transport protocol used for delay tolerant networks when UDP or TCP are not effective. \cite{40565_2004}
+Active since 1982 and managed in collaboration with several national and international space organizations. A 
+contemporary simulation of a current interplanetary network or earth orbit network would follow the protocols and
+standards laid out by CCSDS.
+
+Several public github repos with satellite simulators can be found. 
+$\text{https://github.com/szymonwieloch/DTN/tree/master}$
+$\text{https://goodingc.github.io/ipn_sim/web-app/}$. Both of these limit themselves to earth orbit simulations.
+
+\section{Interplanetary Internet and Background}
+
+An interplanetary network is simply a network which where the nodes are distributed across the surface or orbits of 
+different planets. In the year of 2024 the current interplanetary is limited to Mars, Earth, the Moon, and satellites 
+orbiting the sun. The interplanetary network is also expanded to deep space satellites, like the voyager probes. A sample 
+message path across the network could be that a message is passed from a on earth station through satellite dish to a 
+earth orbiting satellite, then when it has direct line of sight with the Mars MAVEN sends the message. MAVEN then relays the 
+message to a Mars rover like Odyssey. This description hides the complexity which arises in an interplanetary network.
+
+In a traditional network a majority of the network will happen over a wired connection. This means that the error is relatively 
+low. This is no longer the case in the case of satellites sending messages over potentially millions of kilometers. 
+Atmosphere in the Earth orbit, solar radiation, and dust cause increased error in sending the message between entities. 
+Inverse square law dictates that over these large periods the energy needed to send the message so that it can be 
+read at its destination. Even the doppler effect is needed to be accounted for to tune into radio frequencies on the ISS.
+High error means that sending a message can no longer be assumed to be sent reliably and has to be compensated for. 
+Beyond high error, large distances between entities and the physical limitation that messages cannot travel faster 
+than the speed of light. Light, or any other energy takes 8 minutes to get from the Sun to Earth, and thus for a round trip 
+will take 16 minutes to complete a simple handshake, given that the messages were sent correctly.
+
+Other than physical limitations a interplanetary network has a rapidly changing topology at anytime. For instance satellites 
+which orbit planets can have their field of view blocked as they move over the horizon. While terrestrial networks also have 
+rapidly changing networks at the edges, but the difference is that the core of the network is stable traditionally. Where 
+depending on the hour, day, year, the core of the interplanetary network can change drastically. To be able to connect to the 
+network there are no well known places to access the network, except for Earth.
+
+To combat these challenges several protocols have been created. The SCPS TP is a transport protocol which is used on the network 
+layer. UDP is used in the lowest reliable connections between nodes, and TCP is used in the most reliable network 
+conditions. When a reliable and efficient method of network transportation is needed, then SCPS TP is used. It is a modified 
+version of TCP which uses a slow start protocol, reduced handshakes, reliable data transfer with reduced confirmations. Other 
+protocols like store and load, and licklider are used to also overcome these problems.
+
+The rest of the paper remains to simulate these unique challenges, as well as see in what conditions do the current protocols 
+work.
 

### 2024-04-07 network_sim.py
make tcp work

@@ -1,28 +1,16 @@
-"""network_sim
-
-run a simulation of interplanetary network
-"""
-
+from enum import Enum
 import json
-import csv
-from dataclasses import dataclass
-from typing import List
 from ns import ns
 from physics_simulation import get_stats
-
-NodeContainer = ns.cppyy.gbl.ns3.NodeContainer
-NetDeviceContainer = ns.cppyy.gbl.ns3.NetDeviceContainer
-Ipv4InterfaceContainer = ns.cppyy.gbl.ns3.Ipv4InterfaceContainer
 
 ns.cppyy.cppdef(
     """
 #include "CPyCppyy/API.h"
 
-using namespace ns3;
-
-// create a NodeContainer with two nodes
-NodeContainer create_node_pair(NodeContainer nodes, int i, int j) {
-    return NodeContainer(nodes.Get(i), nodes.Get(j));
+// There literally is not any other way to do this
+// I hate having to call Python from C++, but c'est la vie
+void cpp_update_topology() {
+    CPyCppyy::Eval("update_topology()");
 }
 
 // set an interface to be down
@@ -39,297 +27,222 @@ void set_up(NodeContainer c, int index, int interface) {
     ipv4->SetUp(interface);
 }
 
-// print the global routing table
-void print_routing_table() {
-    Ipv4GlobalRoutingHelper globalRouting;
-    Ptr<OutputStreamWrapper> routingStream = Create<OutputStreamWrapper> (&std::cout);
-    globalRouting.PrintRoutingTableAllAt (Seconds(0.1), routingStream );
-}
-
-// get the number of devices
-int get_num_devices(NodeContainer c, int index) {
-    Ptr<Node> n = c.Get(index);
-
-    return n->GetNDevices();
-}
-
-// get the index of a netdevice
-int get_netdevice_node_index(NodeContainer c, int node_index, int device_index) {
-    Ptr<Channel> ch = c.Get(node_index)->GetDevice(device_index)->GetChannel();
-    int i1 = ch->GetDevice(0)->GetNode()->GetId();
-    int i2 = ch->GetDevice(1)->GetNode()->GetId();
-
-    return node_index == i1 ? i2 : i1;
-}
-
-// set the channel delay on a channel
-void set_channel_delay(NodeContainer c, int node_index, int device_index, double seconds) {
-    Ptr<Channel> ch = c.Get(node_index)->GetDevice(device_index)->GetChannel();
+// set the delay on a channel
[…]

### 2024-04-07 network_sim.py
remove unnecessary code

@@ -233,16 +233,6 @@ class Network:
         apps_sink.Start(ns.core.Seconds(1.0))
         apps_sink.Stop(ns.core.Seconds(self.simulation_len))
 
-        # install sink on all routers
-        for i in range(self.num_routers):
-            sink = ns.applications.PacketSinkHelper(
-                factory,
-                ns.network.InetSocketAddress(ns.Ipv4Address.GetAny(), port).ConvertTo(),
-            )
-            apps_sink = sink.Install(self.routers.Get(i))
-            apps_sink.Start(ns.core.Seconds(1.0))
-            apps_sink.Stop(ns.core.Seconds(self.simulation_len))
-
 
 network = Network(10000, Protocol.TCP, "Earth", "Mars", simulation_len=60 * 60)
 network.run()

### 2024-04-07 network_sim.py
fix TCP issues

@@ -3,6 +3,14 @@ import json
 from ns import ns
 from physics_simulation import get_stats
 
+# NOTE ------------------------------------------------------------------------
+# Multiply all times output by this by 26 to get the correct time
+# This divider gets around the very limited TTL on sockets
+# This is a workaround to not have to implement my own TCP protocol from scratch
+# and guarantees that message sent across the diameter of Mars's orbit will not
+# time out before they even arrive
+TIME_DIVIDER = 26
+
 ns.cppyy.cppdef(
     """
 #include "CPyCppyy/API.h"
@@ -133,12 +141,15 @@ class Network:
                 else:
                     details = entities[i]["connections"][connections[j]]
                     if details["connected"]:
-                        set_channel_delay(self.routers, i, j, details["trans_time"])
+                        set_channel_delay(
+                            self.routers, i, j, details["trans_time"] / TIME_DIVIDER
+                        )
                         # set_channel_delay(self.routers, i, j, 61.399)
                         set_channel_error(self.routers, i, j, details["error_rate"])
                         set_up(self.routers, i, j)
                     else:
                         set_down(self.routers, i, j)
+        ns.internet.Ipv4GlobalRoutingHelper.RecomputeRoutingTables()
         self.time += self.time_step
 
     def __create_nodes(self):
@@ -184,7 +195,7 @@ class Network:
         p2p = ns.point_to_point.PointToPointHelper()
         p2p.SetDeviceAttribute("DataRate", ns.core.StringValue("10Mbps"))
         p2p.SetChannelAttribute("Delay", ns.core.StringValue("0ms"))
-        # p2p.EnableAsciiAll(self.stream)
+        p2p.EnableAsciiAll(self.stream)
         self.sender_to_router = p2p.Install(
             self.sender.Get(0), self.routers.Get(sender_id)
         )
@@ -234,5 +245,5 @@ class Network:
         apps_sink.Stop(ns.core.Seconds(self.simulation_len))
 
 
-network = Network(10000, Protocol.TCP, "Earth", "Mars", simulation_len=60 * 60)
+network = Network(10000, Protocol.TCP, "Earth", "Mars", simulation_len=60 * 60 * 24)
 network.run()

### 2024-04-08 network_sim.py
new reno

@@ -51,6 +51,14 @@ void set_channel_error(NodeContainer c, int index, int interface, double rate) {
 
     ch->SetAttribute("ReceiveErrorModel", PointerValue(em));
 }
+
+// void turn on new reno
+void new_reno() {
+    Config::SetDefault(
+        "ns3::TcpL4Protocol::SocketType",
+        TypeIdValue(TypeId::LookupByName("ns3::TcpNewReno"))
+    );
+}
     """
 )
 
@@ -73,6 +81,7 @@ def update_topology():
 class Protocol(Enum):
     UDP = 1
     TCP = 2
+    NEW_RENO = 3
 
 
 class Network:
@@ -96,6 +105,9 @@ class Network:
             "ns3::Ipv4GlobalRouting::RespondToInterfaceEvents",
             ns.core.BooleanValue(True),
         )
+        if protocol == Protocol.NEW_RENO:
+            ns.cppyy.gbl.new_reno()
+            self.protocol = Protocol.TCP
 
         # setup tracing
         self.ascii = ns.network.AsciiTraceHelper()
@@ -245,5 +257,5 @@ class Network:
         apps_sink.Stop(ns.core.Seconds(self.simulation_len))
 
 
-network = Network(10000, Protocol.TCP, "Earth", "Mars", simulation_len=60 * 60 * 24)
+network = Network(10000, Protocol.NEW_RENO, "Earth", "Mars", simulation_len=60 * 60 * 24)
 network.run()

### 2024-04-08 network_sim.py
add stats to readme

@@ -1,5 +1,6 @@
 from enum import Enum
 import json
+import sys
 from ns import ns
 from physics_simulation import get_stats
 
@@ -92,7 +93,7 @@ class Network:
         sender_body: str,
         receiver_body: str,
         time_step: int = 60,
-        simulation_len: int = 60 * 60 * 24,
+        simulation_len: int = 60 * 60,
     ) -> None:
         # assign instance variables
         self.time = start_time
@@ -185,7 +186,7 @@ class Network:
 
     def __connect_routers(self):
         p2p = ns.point_to_point.PointToPointHelper()
-        p2p.SetDeviceAttribute("DataRate", ns.core.StringValue("10Mbps"))
+        p2p.SetDeviceAttribute("DataRate", ns.core.StringValue("1Mbps"))
         p2p.SetChannelAttribute("Delay", ns.core.StringValue("10ms"))
         p2p.EnableAsciiAll(self.stream)
 
@@ -240,8 +241,7 @@ class Network:
                 self.router_to_receiver_address.GetAddress(1), port
             ).ConvertTo(),
         )
-        rate = 8 / self.time_step
-        onoff.SetConstantRate(ns.network.DataRate(f"{rate}kbps"))
+        onoff.SetConstantRate(ns.network.DataRate(f"1Mbps"))
         onoff.SetAttribute("PacketSize", ns.core.UintegerValue(1024))
         apps_onoff = onoff.Install(self.sender.Get(0))
         apps_onoff.Start(ns.core.Seconds(1.0))
@@ -257,5 +257,35 @@ class Network:
         apps_sink.Stop(ns.core.Seconds(self.simulation_len))
 
 
-network = Network(10000, Protocol.NEW_RENO, "Earth", "Mars", simulation_len=60 * 60 * 24)
-network.run()
+def main():
+    if len(sys.argv) < 2:
+        print("Protocol not specified")
+        sys.exit(1)
+    elif sys.argv[1] == "UDP":
+        protocol = Protocol.UDP
+    elif sys.argv[1] == "TCP":
+        protocol = Protocol.TCP
+    elif sys.argv[1] == "NewReno":
+        protocol = Protocol.NEW_RENO
+    else:
+        print(f"Protocol {sys.argv[1]} is not supported")
+        sys.exit(1)
+
+    network = Network(10000, protocol, "Earth", "Mars")
+    network.run()
+
+
+if __name__=="__main__":
+    main()
+
+# 3600 seconds
+# ----------------------------
[…]

### 2024-04-08 physics_simulation.py
Added high error mode

@@ -1,5 +1,6 @@
 import math
 import json
+import sys
 
 # Speed of light m/s
 C = 299792458.0
@@ -348,7 +349,22 @@ def get_error_rate(sender, receiver) -> float:
         and "x" in receiver.keys()
         and "y" in receiver.keys()
     )
-    return free_space_path_loss(sender["x"], sender["y"], receiver["x"], receiver["y"])
+
+    # Calculate error rate based on FSPL
+    error_rate = free_space_path_loss(sender["x"], sender["y"], receiver["x"], receiver["y"])
+
+    # Check if we're in 'high_error' mode
+    if 'high_error' in sys.argv:
+        scaling = 10
+
+        # Check if a specific scaling factor was specified
+        if sys.argv.index('high_error') < len(sys.argv)-1:
+            scaling = int(sys.argv[sys.argv.index('high_error')+1])
+
+        # Increase error rate by 10 orders of magnitude, up to 100%
+        error_rate = min(error_rate*(10**scaling) , 1)
+
+    return error_rate 
 
 
 # for testing

### 2024-04-09 physics_simulation.py
Changed high_error mode

@@ -1,6 +1,5 @@
 import math
 import json
-import sys
 
 # Speed of light m/s
 C = 299792458.0
@@ -164,7 +163,7 @@ def get_entity_stats(t, entity_stats, stat_list):
 
 
 # returns information for all entities at time time t
-def get_stats(t: int):
+def get_stats(t: int, high_error: bool = True):
 
     stats = []
 
@@ -189,7 +188,7 @@ def get_stats(t: int):
             entity_stats = get_entity_stats(t, entity, stats)
             stats.append(entity_stats)
 
-    stats = get_connections(stats)
+    stats = get_connections(stats, high_error)
 
     return stats
 
@@ -253,7 +252,7 @@ def point_dist_to_line(
 # "can_connect": entity able to connect to the interplanetary internet
 # "x": x coordinate in space (m)
 # "y": y coordinate in space (m)
-def get_connections(stats):
+def get_connections(stats, high_error):
 
     # If anyone has a better idea that isnt O(n^3) I am listening
     for entity_sending in stats:
@@ -295,7 +294,7 @@ def get_connections(stats):
                         entity_sending["y"],
                     )
                     trans_time = transmission_time(dist)
-                    err_rate = get_error_rate(entity_sending, entity_receiving)
+                    err_rate = get_error_rate(entity_sending, entity_receiving, high_error)
                 else:
                     dist = None
                     trans_time = None
@@ -342,7 +341,7 @@ def free_space_path_loss(x1, y1, x2, y2) -> float:
 
 
 # Get the rate of transmission error between two satelites
-def get_error_rate(sender, receiver) -> float:
+def get_error_rate(sender, receiver, high_error) -> float:
     assert (
         "x" in sender.keys()
         and "y" in sender.keys()
@@ -354,15 +353,9 @@ def get_error_rate(sender, receiver) -> float:
     error_rate = free_space_path_loss(sender["x"], sender["y"], receiver["x"], receiver["y"])
 
     # Check if we're in 'high_error' mode
-    if 'high_error' in sys.argv:
-        scaling = 10
-
-        # Check if a specific scaling factor was specified
-        if sys.argv.index('high_error') < len(sys.argv)-1:
-            scaling = int(sys.argv[sys.argv.index('high_error')+1])
-
+    if high_error:
         # Increase error rate by 10 orders of magnitude, up to 100%
-        error_rate = min(error_rate*(10**scaling) , 1)
+        error_rate = min(error_rate*(10**10) , 1)
 
     return error_rate 
 

### 2024-04-09 paper/paper.tex
wrote about implementing with ns-3

--- 2024-03-31 paper/paper.tex
@@
 \documentclass[a4paper,12pt]{article}
+
+\usepackage{microtype}
+\usepackage{biblatex}
+
+\addbibresource{paper.bib}
 
@@
 \date{\today}
+
+\newcommand{\CC}{C\nolinebreak\hspace{-.05em}\raisebox{.4ex}{\tiny\bf +}\nolinebreak\hspace{-.10em}\raisebox{.4ex}{\tiny\bf +}}
+\def\CC{{C\nolinebreak[4]\hspace{-.05em}\raisebox{.4ex}{\tiny\bf ++}}}
 
@@
 
-In the past decade interest in returning back to space has increased internationally. Many organizations 
-have started sending remote vehicles into space for the first time \cite{(indian space program, spaceX, ect)}
-and there are talks of a permanent lunar base being created. With an increased amount of satellites and potentially 
-permanent bases outside of Earths orbit, the demand on the interplanetary network (IPN) is increased. The IPN has a 
-series of unique challenges such as a high amount of error, constantly changing network topology, and long 
-propagation delay. The solution to these problems is implementing a delay tolerant network. While space organizations 
-have maintained the IPN with a set of standard protocols, these may not be sufficient for an expanding network.
-The simulator proposed shows how an IPN can handle different stages of development of an IPN as well as increased 
-throughput requirements.
+In the past decade interest in returning back to space has increased
+internationally. Many organizations have started sending remote vehicles into
+space for the first time~%\cite{(indian space program, spaceX, ect)}
+and there
+are talks of a permanent lunar base being created. With an increased amount of
+satellites and potentially permanent bases outside of Earths orbit, the demand
+on the interplanetary network (IPN) is increased. The IPN has a series of unique
+challenges such as a high amount of error, constantly changing network topology,
+and long propagation delay. The solution to these problems is implementing a
+delay tolerant network. While space organizations have maintained the IPN with a
+set of standard protocols, these may not be sufficient for an expanding network.
+The simulator proposed shows how an IPN can handle different stages of
+development of an IPN as well as increased throughput requirements.
 
@@
 
-A significant challenge in creating a IPN is the rapidly changing topology. To simulate 
-these rapid changing conditions a physics engine was created to model the orbits of several 
-entities in the solar system. Several assumptions were made in the simulator for the sake of 
-simplicity as the goal of the simulator is not to create a perfect model of the solar system 
-and the entities inside of the solar system, but rather conditions which a IPN would exist in.
+A significant challenge in creating a IPN is the rapidly changing topology. To
+simulate these rapid changing conditions a physics engine was created to model
+the orbits of several entities in the solar system. Several assumptions were
+made in the simulator for the sake of simplicity as the goal of the simulator is
+not to create a perfect model of the solar system and the entities inside of the
+solar system, but rather conditions which a IPN would exist in.
 
-The first assumption made was that all orbits are circular. Realistically, orbits are oval 
-shaped, with the two focal points. A result of this assumption is the elimination of any 
-difference in orbital speed due to Kepler's third law. The orbital radius was chosen by taking
-the average orbital radius of the real life celestial bodies.
-For example Earth's average orbital
-radius is 149.6 billion meters \cite{}, and therefore orbits around the sun in a circular orbit
-of radius 149.6 billion meters.  Another result of this assumption 
-is that orbits are considered to be centered around the body which it orbits. The Mars orbit 
-has a focal point that is further in space, making it more oval shaped than the earth orbit. \cite{} 
-While this could make changes to a 
-simulated network in some very specific situations, the primary behaviour of a solar system 
-is maintained, like retrograde orbits, periodical alignment, predictable behaviour, and 
-recursive orbits of sub orbiting object.
+The first assumption made was that all orbits are circular. Realistically,
+orbits are oval shaped, with the two focal points. A result of this assumption
+is the elimination of any difference in orbital speed due to Kepler's third law.
+The orbital radius was chosen by taking the average orbital radius of the real
+life celestial bodies. For example Earth's average orbital radius is 149.6
+billion meters~\cite{}, and therefore orbits around the sun in a circular orbit
+of radius 149.6 billion meters. Another result of this assumption is that orbits
+are considered to be centered around the body which it orbits. The Mars orbit
+has a focal point that is further in space, making it more oval shaped than the
+earth orbit.~\cite{} While this could make changes to a simulated network in
+some very specific situations, the primary behaviour of a solar system is
+maintained, like retrograde orbits, periodical alignment, predictable behaviour,
+and recursive orbits of sub orbiting object.
 
-All entities are assumed to have a circular shape, with a constant radius. Due to the scale of the 
-objects, both large and small, this has a very minor effect.
+All entities are assumed to have a circular shape, with a constant radius. Due
+to the scale of the objects, both large and small, this has a very minor effect.
 
-The whole physics simulation has been done on a 2D plane. While planetary orbits were not perfectly
-aligned, on the astronomical scale which we are working on, assuming that all their orbits are 
-contained on the same plane will not effect the simulation significantly. Where this could be an issue 
-is for when there are more than two satellites orbiting the same entity, with a relatively low orbit. For
-example satellites orbiting Earth. While there are some complex satellite interactions which are removed 
-it does still simulate satellites having to form a network around the entity it is 
-orbiting, it just requires less satellites then if it needed to form a 3D network.
+The whole physics simulation has been done on a 2D plane. While planetary orbits
+were not perfectly aligned, on the astronomical scale which we are working on,
+assuming that all their orbits are contained on the same plane will not effect
+the simulation significantly. Where this could be an issue is for when there are
+more than two satellites orbiting the same entity, with a relatively low orbit.
+For example satellites orbiting Earth. While there are some complex satellite
+interactions which are removed it does still simulate satellites having to form
+a network around the entity it is orbiting, it just requires less satellites
+then if it needed to form a 3-D network.
 
-Messages between entities were assumed to be sent at the speed of light, as all forms of radio, lasers,
-or other wireless messages all are sent at the speed of light. Since propagation delay is much higher 
-then terrestrial networks, other delays (TODO specify which delays) were ignored as being negligible.
+Messages between entities were assumed to be sent at the speed of light, as all
+forms of radio, lasers, or other wireless messages all are sent at the speed of
+light. Since propagation delay is much higher then terrestrial networks, other
+delays (TODO specify which delays) were ignored as being negligible.
 
-Based on these simplifications, a physical simulation with several goals. To calculate if a line of 
-sight could be made between two communicating entities, that is if a message can be sent, to calculate 
-the error rate when sending a message, and to calculate the message propagation delay for the message to 
-be sent. Line of sight was achieved using simple geometry calculations. (See appendix) All entities are able to block 
-signals if the ray passes through the radius of the entity.
-One of the largest factors contributing to transmission error is the distance  travelled by the electromagnetic signal between transmitting and receiving nodes.
-By generalizing Friis' transmission formula for isotropic emitters and receivers, we can obtain the ratio of signal intensity received to signal intensity transmitted, known as Free Space Path Loss (FSPL).
-We can use FSPL as our transmission error rate, giving the following expression: $Error Rate = (\frac{4 \pi d}{\lambda})^2$ where $d$ is the distance between nodes in metres and $\lambda$ is the wavelength of the electromagnetic wave used in transmission.
-As much of space communications utilizes the Ka-band of frequencies (27-40 GHz) we opted to use a transmission frequency of 30Ghz.
-The transmission wavelength can be then calculated according to $\lambda = \frac{c}{f}$, where $c$ is the speed of light and $f$ is the transmission frequency, giving a transmission wavelength of ~9.99mm.
-Propagation delay is a simple function of distance and the speed of light. The input of the physics engine 
-is a time $t$ from some preset initial conditions which outputs the state of all entities.
+Based on these simplifications, a physical simulation with several goals. To
+calculate if a line of sight could be made between two communicating entities,
+that is if a message can be sent, to calculate the error rate when sending a
+message, and to calculate the message propagation delay for the message to be
+sent. Line of sight was achieved using simple geometry calculations. (See
+appendix) All entities are able to block signals if the ray passes through the
+radius of the entity. One of the largest factors contributing to transmission
+error is the distance travelled by the electromagnetic signal between
+transmitting and receiving nodes. By generalizing Friis' transmission formula
+for isotropic emitters and receivers, we can obtain the ratio of signal
+intensity received to signal intensity transmitted, known as Free Space Path
+Loss (FSPL). We can use FSPL as our transmission error rate, giving the
+following expression: $Error Rate = {(\frac{4 \pi d}{\lambda})}^2$ where $d$ is
+the distance between nodes in metres and $\lambda$ is the wavelength of the
+electromagnetic wave used in transmission. As much of space communications
+utilizes the Ka-band of frequencies (27--40 GHz) we opted to use a transmission
+frequency of 30Ghz. The transmission wavelength can be then calculated according
+to $\lambda = \frac{c}{f}$, where $c$ is the speed of light and $f$ is the
+transmission frequency, giving a transmission wavelength of about 9.99~mm.
+Propagation delay is a simple function of distance and the speed of light. The
+input of the physics engine is a time $t$ from some preset initial conditions
+which outputs the state of all entities.
+
+\subsection{Network Simulation}
+
+We used NS-3~\cite{ns-3} to implement a simulation of this network. Our final design
+consisted of a set of routers representing the different entities in our physics
+simulation which were each linked to every other router using a point-to-point
+channel. Essentially, this allowed us to have a wired network with configurable
+delay and error along with the ability to take certain channels down if they are
+obstructed that we could make behave like an interplanetary network would
+without having to deal with the complexities of configuring wireless
+communication. In addition to the mesh of routers, we added two individual nodes
+to the network with zero delay that can be connected to any one router in the
+network. One of these two nodes acts as a sender and the other one acts as a
+receiver. This configuration of sender and receiver as separate nodes made it
+easier to decouple the applications from the network topology and allowed us to
+programatically create arbitrary topologies.
+
+NS-3 is a useful library, but we did run into a number of problems over the
+course of our implementation of our simulation.
+The first problem came when we first tried to compile the library itself and ran into a bug in the build script that caused all executables to fail to run if there was the word ``scratch'' anywhere in the path due to a specific subdirectory of NS-3 which also happens to be called ``scratch''.
+When first testing it out, one of us had put NS-3 into a directory that happened to contain that word.
+We made a pull request to fix this issue and it was merged by the NS-3 maintainers which allowed us to continue.
+
+The next issue we ran into was that we had decided to use the Python bindings to
+allow us to easily use Python for the physics simulation. This was a mistake.
+Although NS-3 does have Python bindings that do make it easier to run it as a
+script after writing, there is no documentation at all for the Python bindings
+and some basic features such as setting the error rate on a channel or
+scheduling a task to run during the simulation have no Python support and must
+use inline \CC. The lack of support for task scheduling in Python is
+particularly bad as it forced us to schedule a \CC{} function which itself
+called Python code. It also forced us to use global state as \texttt{CPyCpyy}
+does not have support for arbitrary function parameters when calling Python code
+as far as we could tell (this feature is not documented and we only figured out
+about it because the few Python examples in the NS-3 source code that do use
+scheduling also call Python from \CC{} with this method).
+
+Before we settled on using point-to-point to emulate a wireless network, we
+attempted to use the existing wireless networking features of NS-3. These
+features are very powerful and even have a positioning and velocity system for
+nodes in the network which would be very useful. We ran into a lot of issues
+with this system though. Firstly, there only seems to be real support for Wi-Fi
+which is not particularly useful for long distance unlike radio. This meant that
+we would have to greatly decrease the distances that we were simulating. This is
+fine and would still allow a perfectly reasonable simulation so long as we
+increased the distances again in our computations while analyzing the data. The
+other problem that we ran into is the fact that the wireless networking would
+quickly require us to go well outside the scope of our project. Determining the
+gain and energy required to configure our satellites to send and receive
+messages correctly is more of an engineering problem that we were not interested
+in solving.
+
+After settling on using point-to-point, UDP was easy to get working but TCP
+proved to be very challenging. Our initial implementation of our topology seemed
+to have an issue where messages could be routed in one direction but not the
+other. This meant that when TCP tried to establish a connection it would always
+fail. It still is not entirely clear what was configured incorrectly, but after
+changing how we were creating channels, TCP was finally working. We ran into
+more issues with TCP though, when we started adding delay to the channels. It
+turns out that NS-3 sockets have a rather low time-to-live and there do not seem
+to be any global options for configuring sockets. We were left with two options:
+implement TCP from scratch using custom-built sockets for each node in the
+network or ensure that the messages that we do send have lower delay than the
+TTL for our sockets. We opted for the latter. To make sure that TCP messages
+could successfully ACK, we found a value that we could divide all times by to
+make sure that almost every message that could feasibly be sent would be
+received before its TTL expired. This does not change any distances or error
+rates and effectively only changes the unit we use for time from $s$ to
+$\frac{s}{26}$ which has an easy conversion back to $s$. Finally, we implemented
+a version of our program using the New Reno variant of TCP. Thankfully, NS-3
+made this easy for us. It worked perfectly after assigning a configuration value
+before setting up our communication channels.
+
+\printbibliography{}
 

### 2024-04-09 paper/paper.tex

@@
 development of an IPN as well as increased throughput requirements.
+Before these technological changes can be implemented simulations need to be created to 
+lower the risk of failure given the high cost of failure. Often satellites will need to be in operation for years 
+or even decades. The hardware and software in satellites can often not be updated very well. Thus it will be very 
+relevant to investigate these issues well before they are needed.
 
 \section{Previous Work}
+
+In their 2017 Masters Thesis, Md Monjurul Islam Khan \cite{Khan2017} from University of Manitoba wrote on how 
+to create a satellite network
+simulation based on ns3. They used a point to point ns3 module to create a network of 2 levels of orbiting 
+satellites and fixed points on the earth which acts as uplinks to the network. The paper analyzed TCP 
+HighSpeed against TCP Vegas slow start. However this paper stops at expanding the network beyond earth orbit, 
+and rather focuses on Earth orbit communication.
+
+SNS3 is a simulation library which uses NS3 as a basis to create a framework which to build simulations for 
+geostationary satellites in the european continent. It does not expand to an interplanetary network.
+\cite{10.4108/icst.simutools.2014.254631}
+
+For non simulation work done on the interplanetary internet, the leading authority is Consultative Committee for 
+Space Data Systems (CCSDS). \cite{CCSDS.org} They maintain and publish the standards for interplanetary internet, including 
+the SCPS TP, a reliable transport protocol used for delay tolerant networks when UDP or TCP are not effective. \cite{40565_2004}
+Active since 1982 and managed in collaboration with several national and international space organizations. A 
+contemporary simulation of a current interplanetary network or earth orbit network would follow the protocols and
+standards laid out by CCSDS.
+
+Several public github repos with satellite simulators can be found. 
+$\text{https://github.com/szymonwieloch/DTN/tree/master}$
+$\text{https://goodingc.github.io/ipn_sim/web-app/}$. Both of these limit themselves to earth orbit simulations.
+
+\section{Interplanetary Internet and Background}
+
+An interplanetary network is simply a network which where the nodes are distributed across the surface or orbits of 
+different planets. In the year of 2024 the current interplanetary is limited to Mars, Earth, the Moon, and satellites 
+orbiting the sun. The interplanetary network is also expanded to deep space satellites, like the voyager probes. A sample 
+message path across the network could be that a message is passed from a on earth station through satellite dish to a 
+earth orbiting satellite, then when it has direct line of sight with the Mars MAVEN sends the message. MAVEN then relays the 
+message to a Mars rover like Odyssey. This description hides the complexity which arises in an interplanetary network.
+
+In a traditional network a majority of the network will happen over a wired connection. This means that the error is relatively 
+low. This is no longer the case in the case of satellites sending messages over potentially millions of kilometers. 
+Atmosphere in the Earth orbit, solar radiation, and dust cause increased error in sending the message between entities. 
+Inverse square law dictates that over these large periods the energy needed to send the message so that it can be 
+read at its destination. Even the doppler effect is needed to be accounted for to tune into radio frequencies on the ISS.
+High error means that sending a message can no longer be assumed to be sent reliably and has to be compensated for. 
+Beyond high error, large distances between entities and the physical limitation that messages cannot travel faster 
+than the speed of light. Light, or any other energy takes 8 minutes to get from the Sun to Earth, and thus for a round trip 
+will take 16 minutes to complete a simple handshake, given that the messages were sent correctly.
+
+Other than physical limitations a interplanetary network has a rapidly changing topology at anytime. For instance satellites 
+which orbit planets can have their field of view blocked as they move over the horizon. While terrestrial networks also have 
+rapidly changing networks at the edges, but the difference is that the core of the network is stable traditionally. Where 
+depending on the hour, day, year, the core of the interplanetary network can change drastically. To be able to connect to the 
+network there are no well known places to access the network, except for Earth.
+
+To combat these challenges several protocols have been created. The SCPS TP is a transport protocol which is used on the network 
+layer. UDP is used in the lowest reliable connections between nodes, and TCP is used in the most reliable network 
+conditions. When a reliable and efficient method of network transportation is needed, then SCPS TP is used. It is a modified 
+version of TCP which uses a slow start protocol, reduced handshakes, reliable data transfer with reduced confirmations. Other 
+protocols like store and load, and licklider are used to also overcome these problems.
+
+The rest of the paper remains to simulate these unique challenges, as well as see in what conditions do the current protocols 
+work.
 
@@
 NS-3 is a useful library, but we did run into a number of problems over the
-course of our implementation of our simulation.
-The first problem came when we first tried to compile the library itself and ran into a bug in the build script that caused all executables to fail to run if there was the word ``scratch'' anywhere in the path due to a specific subdirectory of NS-3 which also happens to be called ``scratch''.
-When first testing it out, one of us had put NS-3 into a directory that happened to contain that word.
-We made a pull request to fix this issue and it was merged by the NS-3 maintainers which allowed us to continue.
+course of our implementation of our simulation. The first problem came when we
+first tried to compile the library itself and ran into a bug in the build script
+that caused all executables to fail to run if there was the word ``scratch''
+anywhere in the path due to a specific subdirectory of NS-3 which also happens
+to be called ``scratch''. When first testing it out, one of us had put NS-3 into
+a directory that happened to contain that word. We made a pull request to fix
+this issue and it was merged by the NS-3 maintainers which allowed us to
+continue.
 

### 2024-04-09 paper/paper.tex
fix compilation and formatting issues

@@
 \usepackage{biblatex}
+\usepackage{hyperref}
 
@@
 space for the first time~%\cite{(indian space program, spaceX, ect)}
-and there
-are talks of a permanent lunar base being created. With an increased amount of
-satellites and potentially permanent bases outside of Earths orbit, the demand
-on the interplanetary network (IPN) is increased. The IPN has a series of unique
-challenges such as a high amount of error, constantly changing network topology,
-and long propagation delay. The solution to these problems is implementing a
-delay tolerant network. While space organizations have maintained the IPN with a
-set of standard protocols, these may not be sufficient for an expanding network.
-The simulator proposed shows how an IPN can handle different stages of
-development of an IPN as well as increased throughput requirements.
-Before these technological changes can be implemented simulations need to be created to 
-lower the risk of failure given the high cost of failure. Often satellites will need to be in operation for years 
-or even decades. The hardware and software in satellites can often not be updated very well. Thus it will be very 
-relevant to investigate these issues well before they are needed.
+and there are talks of a permanent lunar base being created. With an increased
+amount of satellites and potentially permanent bases outside of Earths orbit,
+the demand on the interplanetary network (IPN) is increased. The IPN has a
+series of unique challenges such as a high amount of error, constantly changing
+network topology, and long propagation delay. The solution to these problems is
+implementing a delay tolerant network. While space organizations have maintained
+the IPN with a set of standard protocols, these may not be sufficient for an
+expanding network. The simulator proposed shows how an IPN can handle different
+stages of development of an IPN as well as increased throughput requirements.
+Before these technological changes can be implemented simulations need to be
+created to lower the risk of failure given the high cost of failure. Often
+satellites will need to be in operation for years or even decades. The hardware
+and software in satellites can often not be updated very well. Thus it will be
+very relevant to investigate these issues well before they are needed.
 
@@
 
-In their 2017 Masters Thesis, Md Monjurul Islam Khan \cite{Khan2017} from University of Manitoba wrote on how 
-to create a satellite network
-simulation based on ns3. They used a point to point ns3 module to create a network of 2 levels of orbiting 
-satellites and fixed points on the earth which acts as uplinks to the network. The paper analyzed TCP 
-HighSpeed against TCP Vegas slow start. However this paper stops at expanding the network beyond earth orbit, 
-and rather focuses on Earth orbit communication.
+In their 2017 Masters Thesis, Md Monjurul Islam Khan~\cite{Khan2017} from
+University of Manitoba wrote on how to create a satellite network simulation
+based on ns3. They used a point to point ns3 module to create a network of 2
+levels of orbiting satellites and fixed points on the earth which acts as
+uplinks to the network. The paper analyzed TCP HighSpeed against TCP Vegas slow
+start. However this paper stops at expanding the network beyond earth orbit, and
+rather focuses on Earth orbit communication.
 
-SNS3 is a simulation library which uses NS3 as a basis to create a framework which to build simulations for 
-geostationary satellites in the european continent. It does not expand to an interplanetary network.
-\cite{10.4108/icst.simutools.2014.254631}
+SNS3 is a simulation library which uses NS3 as a basis to create a framework
+which to build simulations for geostationary satellites in the European
+continent. It does not expand to an interplanetary network~\cite{Puttonen2014}.
 
-For non simulation work done on the interplanetary internet, the leading authority is Consultative Committee for 
-Space Data Systems (CCSDS). \cite{CCSDS.org} They maintain and publish the standards for interplanetary internet, including 
-the SCPS TP, a reliable transport protocol used for delay tolerant networks when UDP or TCP are not effective. \cite{40565_2004}
-Active since 1982 and managed in collaboration with several national and international space organizations. A 
-contemporary simulation of a current interplanetary network or earth orbit network would follow the protocols and
-standards laid out by CCSDS.
+For non simulation work done on the interplanetary internet, the leading
+authority is Consultative Committee for Space Data Systems
+(CCSDS)~\cite{CCSDS.org}. They maintain and publish the standards for
+interplanetary internet, including the SCPS TP, a reliable transport protocol
+used for delay tolerant networks when UDP or TCP are not
+effective.~\cite{Keith2004} Active since 1982 and managed in collaboration with
+several national and international space organizations. A contemporary
+simulation of a current interplanetary network or earth orbit network would
+follow the protocols and standards laid out by CCSDS.
 
-Several public github repos with satellite simulators can be found. 
-$\text{https://github.com/szymonwieloch/DTN/tree/master}$
-$\text{https://goodingc.github.io/ipn_sim/web-app/}$. Both of these limit themselves to earth orbit simulations.
+Several public GitHub repositories with satellite simulators can be
+found, for instance:
+\begin{itemize}
+\item\url{https://github.com/szymonwieloch/DTN/tree/master}
+\item\url{https://goodingc.github.io/ipn_sim/web-app/}
+\end{itemize}
+Both of these limit themselves to earth orbit simulations.
 
@@
 
-An interplanetary network is simply a network which where the nodes are distributed across the surface or orbits of 
-different planets. In the year of 2024 the current interplanetary is limited to Mars, Earth, the Moon, and satellites 
-orbiting the sun. The interplanetary network is also expanded to deep space satellites, like the voyager probes. A sample 
-message path across the network could be that a message is passed from a on earth station through satellite dish to a 
-earth orbiting satellite, then when it has direct line of sight with the Mars MAVEN sends the message. MAVEN then relays the 
-message to a Mars rover like Odyssey. This description hides the complexity which arises in an interplanetary network.
+An interplanetary network is simply a network which where the nodes are
+distributed across the surface or orbits of different planets. In the year of
+2024 the current interplanetary is limited to Mars, Earth, the Moon, and
+satellites orbiting the sun. The interplanetary network is also expanded to deep
+space satellites, like the voyager probes. A sample message path across the
+network could be that a message is passed from a on earth station through
+satellite dish to a earth orbiting satellite, then when it has direct line of
+sight with the Mars MAVEN sends the message. MAVEN then relays the message to a
+Mars rover like Odyssey. This description hides the complexity which arises in
+an interplanetary network.
 
-In a traditional network a majority of the network will happen over a wired connection. This means that the error is relatively 
-low. This is no longer the case in the case of satellites sending messages over potentially millions of kilometers. 
-Atmosphere in the Earth orbit, solar radiation, and dust cause increased error in sending the message between entities. 
-Inverse square law dictates that over these large periods the energy needed to send the message so that it can be 
-read at its destination. Even the doppler effect is needed to be accounted for to tune into radio frequencies on the ISS.
-High error means that sending a message can no longer be assumed to be sent reliably and has to be compensated for. 
-Beyond high error, large distances between entities and the physical limitation that messages cannot travel faster 
-than the speed of light. Light, or any other energy takes 8 minutes to get from the Sun to Earth, and thus for a round trip 
-will take 16 minutes to complete a simple handshake, given that the messages were sent correctly.
+In a traditional network a majority of the network will happen over a wired
+connection. This means that the error is relatively low. This is no longer the
+case in the case of satellites sending messages over potentially millions of
+kilometers. Atmosphere in the Earth orbit, solar radiation, and dust cause
+increased error in sending the message between entities. Inverse square law
+dictates that over these large periods the energy needed to send the message so
+that it can be read at its destination. Even the Doppler effect is needed to be
+accounted for to tune into radio frequencies on the ISS. High error means that
+sending a message can no longer be assumed to be sent reliably and has to be
+compensated for. Beyond high error, large distances between entities and the
+physical limitation that messages cannot travel faster than the speed of light.
+Light, or any other energy takes 8 minutes to get from the Sun to Earth, and
+thus for a round trip will take 16 minutes to complete a simple handshake, given
+that the messages were sent correctly.
 
-Other than physical limitations a interplanetary network has a rapidly changing topology at anytime. For instance satellites 
-which orbit planets can have their field of view blocked as they move over the horizon. While terrestrial networks also have 
-rapidly changing networks at the edges, but the difference is that the core of the network is stable traditionally. Where 
-depending on the hour, day, year, the core of the interplanetary network can change drastically. To be able to connect to the 
-network there are no well known places to access the network, except for Earth.
+Other than physical limitations a interplanetary network has a rapidly changing
+topology at anytime. For instance satellites which orbit planets can have their
+field of view blocked as they move over the horizon. While terrestrial networks
+also have rapidly changing networks at the edges, but the difference is that the
+core of the network is stable traditionally. Where depending on the hour, day,
+year, the core of the interplanetary network can change drastically. To be able
+to connect to the network there are no well known places to access the network,
+except for Earth.
 
-To combat these challenges several protocols have been created. The SCPS TP is a transport protocol which is used on the network 
-layer. UDP is used in the lowest reliable connections between nodes, and TCP is used in the most reliable network 
-conditions. When a reliable and efficient method of network transportation is needed, then SCPS TP is used. It is a modified 
-version of TCP which uses a slow start protocol, reduced handshakes, reliable data transfer with reduced confirmations. Other 
-protocols like store and load, and licklider are used to also overcome these problems.
+To combat these challenges several protocols have been created. The SCPS TP is a
+transport protocol which is used on the network layer. UDP is used in the lowest
+reliable connections between nodes, and TCP is used in the most reliable network
+conditions. When a reliable and efficient method of network transportation is
+needed, then SCPS TP is used. It is a modified version of TCP which uses a slow
+start protocol, reduced handshakes, reliable data transfer with reduced
+confirmations. Other protocols like store and load, and Licklider are used to
+also overcome these problems.
 
-The rest of the paper remains to simulate these unique challenges, as well as see in what conditions do the current protocols 
-work.
+The rest of the paper remains to simulate these unique challenges, as well as
+see in what conditions do the current protocols work.
 

### 2024-04-09 process_data.py
added stats and graphs

@@ -1,5 +1,8 @@
+import math
 import re
 import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
 
 
 # take trace file and turn it into a dataframe
@@ -25,7 +28,9 @@ def process_trace():
 
             id = re.compile("(?: id )(.*?(?= ))").search(line).group(1)
 
-            processed_data.append([type, time, node, interface, id])
+            bytes = re.compile("(?: length: )(.*?(?= ))").search(line).group(1)
+
+            processed_data.append([type, time, node, interface, bytes, id])
 
         except Exception as e:
             print("exception encountered")
@@ -33,7 +38,7 @@ def process_trace():
             print(e)
 
     
-    df = pd.DataFrame(processed_data, columns=["type", "time", "node1", "interface", "id"])
+    df = pd.DataFrame(processed_data, columns=["type", "time", "node1", "interface", "bytes", "id"])
     return df
 
 
@@ -50,7 +55,7 @@ def create_stats_df():
 
     df["node2"] = df["Output Node"]
     
-    df = df[["node1", "interface", "node2", "type", "time", "id"]]
+    df = df[["node1", "interface", "node2", "type", "time", "bytes", "id"]]
 
     df = df.drop_duplicates()
 
@@ -63,12 +68,106 @@ def create_stats_df():
     df = df_queue.merge(df_dequeue, how="left", on="id")
     df = df.merge(df_type, how="left", on="id")
 
-    df[["node_sending", "node_receiving", "interface", "time_queued", "time_dequeued", "time_received", "id"]] = df[["node1_x", "node2_x", "interface_x", "time_x", "time_y", "time", "id"]]
+    print(df)
 
-    df = df[["node_sending", "node_receiving", "interface", "time_queued", "time_dequeued", "time_received", "id"]]
+    df[["node_sending", "node_receiving", "interface", "time_queued", "time_dequeued", "time_received", "bytes", "id"]] = df[["node1_x", "node2_x", "interface_x", "time_x", "time_y", "time", "bytes_x", "id"]]
+
+    df = df[["node_sending", "node_receiving", "interface", "time_queued", "time_dequeued", "time_received", "bytes", "id"]]
 
     return df
 
 
+def expand_dataframe(df):
+    # columns: "node_sending", "node_receiving", "interface", "time_queued", "time_dequeued", "time_received", "id"
+
+    df[["time_queued", "time_dequeued", "time_received"]] = df[["time_queued", "time_dequeued", "time_received"]].astype(float)
+
[…]

### 2024-04-09 process_data.py
styled graphs

@@ -113,6 +113,7 @@ def calculate_statistics(df):
 
 
 def graph_stats(df):
+    plt.style.use('ggplot')
     # graph the average message time in transit
     # get max and min time
     min = math.floor(df["time_queued"].min())
@@ -120,19 +121,23 @@ def graph_stats(df):
 
     # create time splits
     # 1000 seconds time step
-    time_step = np.arange(min, max, 1000)
-    x = np.arange(min + 500, max - 500, 1000)
+    time_delta = 1000
+    time_step = np.arange(min, max, time_delta)
+    x = np.arange(min + time_delta/2, max - time_delta/2, time_delta)
 
     grouped_df = df.groupby(pd.cut(df["time_dequeued"], time_step, include_lowest=True)).mean()
 
-    print("time step", len(time_step), min, max)
-    print(len(grouped_df))
-
     plt.plot(x, grouped_df["total_time"])
+    plt.xlabel("time (ms)")
+    plt.ylabel("total message time (ms)")
+    plt.title("Average message time in transit over time")
     plt.show()
     plt.clf()
 
     plt.plot(x, grouped_df["bytes"])
+    plt.xlabel("time (ms)")
+    plt.ylabel("bytes")
+    plt.title("Average message size")
     plt.show()
     plt.clf()
 
@@ -140,12 +145,8 @@ def graph_stats(df):
 def get_graphs():
     print("Create graphs")
 
-    # try:
     df = pd.read_csv("data/message_stats.csv")
     graph_stats(df)
-    # except Exception as e:
-    #     print("ran into error, likely could not find file")
-    #     print(e)
 
 
 def get_statistics():

### 2024-04-10 paper/paper.tex
add abstract

@@
 \begin{abstract}
-    TODO insert abstract here
+  This paper discusses interplanetary networks, the challenges that can arise
+  from them that make them a unique challenge compared to shorter-range computer
+  networks, and the protocols that can be used in this type of network. In
+  addition, we look at how one can use conventional network simulators to test
+  interplanetary network implementations. We used the network simulator
+  NS-3~\cite{ns-3} to test various transport layer protocols against each other
+  in a network topology that we designed in the network simulator to emulate
+  long-distance communications. Using this network simulation, we were able to
+  compare UDP, TCP, and TCP New Reno protocols against each other to see how
+  they each performed across such large distances with high delays.
 \end{abstract}

### 2024-04-10 paper/paper.tex
Added citations

@@
 
-Messages between entities were assumed to be sent at the speed of light, as all
-forms of radio, lasers, or other wireless messages all are sent at the speed of
-light. Since propagation delay is much higher then terrestrial networks, other
-delays (TODO specify which delays) were ignored as being negligible.
+Messages between entities were assumed to be sent at the speed of light in a vacuum, as all
+forms of radio, lasers, or other wireless messages are forms of electromagnetic radiation,
+and thus travel at the speed of
+light. Since the propagation delay in an IPN is much higher than in terrestrial networks, other
+delays such as queuing or processing delays were ignored as being negligible.
 
@@
 error is the distance travelled by the electromagnetic signal between
-transmitting and receiving nodes. By generalizing Friis' transmission formula
+transmitting and receiving nodes. By generalizing Friis' transmission formula\cite{Friis}
 for isotropic emitters and receivers, we can obtain the ratio of signal
@@
 electromagnetic wave used in transmission. As much of space communications
-utilizes the Ka-band of frequencies (27--40 GHz) we opted to use a transmission
+utilizes the Ka-band of frequencies (27--40 GHz)\cite{Morabito_Hastrup} we opted to use a transmission
 frequency of 30Ghz. The transmission wavelength can be then calculated according

### 2024-04-10 paper/paper.tex
add missing citations

@@
 life celestial bodies. For example Earth's average orbital radius is 149.6
-billion meters~\cite{}, and therefore orbits around the sun in a circular orbit
-of radius 149.6 billion meters. Another result of this assumption is that orbits
-are considered to be centered around the body which it orbits. The Mars orbit
-has a focal point that is further in space, making it more oval shaped than the
-earth orbit.~\cite{} While this could make changes to a simulated network in
-some very specific situations, the primary behaviour of a solar system is
-maintained, like retrograde orbits, periodical alignment, predictable behaviour,
-and recursive orbits of sub orbiting object.
+billion meters~\cite{nasa_earth}, and therefore orbits around the sun in a
+circular orbit of radius 149.6 billion meters. Another result of this assumption
+is that orbits are considered to be centered around the body which it orbits.
+The Mars orbit has a focal point that is further in space, making it more oval
+shaped than the earth orbit.~\cite{nasa_mars} While this could make changes to a
+simulated network in some very specific situations, the primary behaviour of a
+solar system is maintained, like retrograde orbits, periodical alignment,
+predictable behaviour, and recursive orbits of sub orbiting object.
 
@@
 
-Messages between entities were assumed to be sent at the speed of light in a vacuum, as all
-forms of radio, lasers, or other wireless messages are forms of electromagnetic radiation,
-and thus travel at the speed of
-light. Since the propagation delay in an IPN is much higher than in terrestrial networks, other
+Messages between entities were assumed to be sent at the speed of light in a
+vacuum, as all forms of radio, lasers, or other wireless messages are forms of
+electromagnetic radiation, and thus travel at the speed of light. Since the
+propagation delay in an IPN is much higher than in terrestrial networks, other
 delays such as queuing or processing delays were ignored as being negligible.
@@
 error is the distance travelled by the electromagnetic signal between
-transmitting and receiving nodes. By generalizing Friis' transmission formula\cite{Friis}
-for isotropic emitters and receivers, we can obtain the ratio of signal
-intensity received to signal intensity transmitted, known as Free Space Path
-Loss (FSPL). We can use FSPL as our transmission error rate, giving the
-following expression: $Error Rate = {(\frac{4 \pi d}{\lambda})}^2$ where $d$ is
-the distance between nodes in metres and $\lambda$ is the wavelength of the
-electromagnetic wave used in transmission. As much of space communications
-utilizes the Ka-band of frequencies (27--40 GHz)\cite{Morabito_Hastrup} we opted to use a transmission
-frequency of 30Ghz. The transmission wavelength can be then calculated according
-to $\lambda = \frac{c}{f}$, where $c$ is the speed of light and $f$ is the
+transmitting and receiving nodes. By generalizing Friis' transmission
+formula\cite{Friis} for isotropic emitters and receivers, we can obtain the
+ratio of signal intensity received to signal intensity transmitted, known as
+Free Space Path Loss (FSPL). We can use FSPL as our transmission error rate,
+giving the following expression: $Error Rate = {(\frac{4 \pi d}{\lambda})}^2$
+where $d$ is the distance between nodes in metres and $\lambda$ is the
+wavelength of the electromagnetic wave used in transmission. As much of space
+communications utilizes the Ka-band of frequencies (27--40
+GHz)\cite{Morabito_Hastrup} we opted to use a transmission frequency of 30Ghz.
+The transmission wavelength can be then calculated according to
+$\lambda = \frac{c}{f}$, where $c$ is the speed of light and $f$ is the
 transmission frequency, giving a transmission wavelength of about 9.99~mm.
@@
 
-We used NS-3~\cite{ns-3} to implement a simulation of this network. Our final design
-consisted of a set of routers representing the different entities in our physics
-simulation which were each linked to every other router using a point-to-point
-channel. Essentially, this allowed us to have a wired network with configurable
-delay and error along with the ability to take certain channels down if they are
-obstructed that we could make behave like an interplanetary network would
-without having to deal with the complexities of configuring wireless
-communication. In addition to the mesh of routers, we added two individual nodes
-to the network with zero delay that can be connected to any one router in the
-network. One of these two nodes acts as a sender and the other one acts as a
-receiver. This configuration of sender and receiver as separate nodes made it
-easier to decouple the applications from the network topology and allowed us to
-programatically create arbitrary topologies.
+We used NS-3~\cite{ns-3} to implement a simulation of this network. Our final
+design consisted of a set of routers representing the different entities in our
+physics simulation which were each linked to every other router using a
+point-to-point channel. Essentially, this allowed us to have a wired network
+with configurable delay and error along with the ability to take certain
+channels down if they are obstructed that we could make behave like an
+interplanetary network would without having to deal with the complexities of
+configuring wireless communication. In addition to the mesh of routers, we added
+two individual nodes to the network with zero delay that can be connected to any
+one router in the network. One of these two nodes acts as a sender and the other
+one acts as a receiver. This configuration of sender and receiver as separate
+nodes made it easier to decouple the applications from the network topology and
+allowed us to programatically create arbitrary topologies.
 

### 2024-04-10 draw_orbits.py
orbit gifs

@@ -28,7 +28,7 @@ def animate(i):
     t =  2 * SECONDS_IN_DAY * i  
 
     # slow down timer to see moon orbit earth
-    # t = 3600 * i
+    t = 360 * i
 
     x_data = []
     y_data = []
@@ -43,8 +43,8 @@ def animate(i):
         y_data.append(datum["y"])
 
     # uncomment to see moon orbit earth
-    # ax.set_xlim([data[1]["x"] - EARTH_ORBIT_RAD, data[1]["x"] + EARTH_ORBIT_RAD])
-    # ax.set_ylim([data[1]["y"] - EARTH_ORBIT_RAD, data[1]["y"] + EARTH_ORBIT_RAD]) 
+    ax.set_xlim([data[1]["x"] - EARTH_ORBIT_RAD, data[1]["x"] + EARTH_ORBIT_RAD])
+    ax.set_ylim([data[1]["y"] - EARTH_ORBIT_RAD, data[1]["y"] + EARTH_ORBIT_RAD]) 
 
     # add it to animation
     scatter.set_offsets(np.column_stack((x_data, y_data)))  
@@ -53,8 +53,11 @@ def animate(i):
 
 # bootstrap for animated python file
 def animated():
-    anim = animation.FuncAnimation(fig, animate, frames = 5000, interval = 20, blit = True)
-    plt.show()
+    anim = animation.FuncAnimation(fig, animate, frames = 500, interval = 20, blit = True)
+    # plt.show()
+
+    writergif = animation.PillowWriter(fps=30) 
+    anim.save("data/orbit_earth.gif", writer=writergif)
 
 
 # non animated daily iteration through drawing

### 2024-04-10 paper/paper.tex
added data section

@@
 \usepackage{hyperref}
+\usepackage{graphicx}
+\usepackage{amsmath}
 
@@
   from them that make them a unique challenge compared to shorter-range computer
-  networks, and the protocols that can be used in this type of network. In
+  networks, and how our current network protocols can be applied in this application. In
   addition, we look at how one can use conventional network simulators to test
   interplanetary network implementations. We used the network simulator
-  NS-3~\cite{ns-3} to test various transport layer protocols against each other
+  NS-3 ~\cite{ns-3} to test various transport layer protocols against each other
   in a network topology that we designed in the network simulator to emulate
@@
   compare UDP, TCP, and TCP New Reno protocols against each other to see how
-  they each performed across such large distances with high delays.
+  they each performed across such large distances with high delays. UDP preformed the 
+  best with New Reno and TCP closely paired together.
 \end{abstract}
@@
 University of Manitoba wrote on how to create a satellite network simulation
-based on ns3. They used a point to point ns3 module to create a network of 2
+based on NS-3. They used a point to point NS-3 module to create a network of 2
 levels of orbiting satellites and fixed points on the earth which acts as
@@
 
-SNS3 is a simulation library which uses NS3 as a basis to create a framework
+SNS3 is a simulation library which uses NS-3 as a basis to create a framework
 which to build simulations for geostationary satellites in the European
@@
 An interplanetary network is simply a network which where the nodes are
-distributed across the surface or orbits of different planets. In the year of
-2024 the current interplanetary is limited to Mars, Earth, the Moon, and
+distributed across the surface or in the orbits of different planets or other celestial objects.
+In the year of
+2024 the current interplanetary is limited to Mars, Earth, Earth orbit, Moon orbit, and
 satellites orbiting the sun. The interplanetary network is also expanded to deep
-space satellites, like the voyager probes. A sample message path across the
-network could be that a message is passed from a on earth station through
-satellite dish to a earth orbiting satellite, then when it has direct line of
-sight with the Mars MAVEN sends the message. MAVEN then relays the message to a
+space satellites, like the Voyager probes. A sample message path across the
+network could be that a message is passed from a large network antenna on earth
+to a earth orbiting satellite, then when it has direct line of
+sight with MAVEN (a satellite orbiting mars) it sends the message. MAVEN then relays the message to a
 Mars rover like Odyssey. This description hides the complexity which arises in
-an interplanetary network.
+an interplanetary network which may seem like a extremely simple network.
 
-In a traditional network a majority of the network will happen over a wired
-connection. This means that the error is relatively low. This is no longer the
+In a traditional network a majority of the network traffic will occur over a wired
+connection, especially for transfers of large amounts of data.
+This means that the error is relatively low. This is no longer the
 case in the case of satellites sending messages over potentially millions of
@@
 that it can be read at its destination. Even the Doppler effect is needed to be
-accounted for to tune into radio frequencies on the ISS. High error means that
+accounted for to tune into radio frequencies on the ISS. ~\cite{qslCompensatingDoppler} High error means that
 sending a message can no longer be assumed to be sent reliably and has to be
@@
 physical limitation that messages cannot travel faster than the speed of light.
-Light, or any other energy takes 8 minutes to get from the Sun to Earth, and
+Light, or any other energy takes approximately 8 minutes to get from the Sun to Earth, ~\cite{nasa_earth} and
 thus for a round trip will take 16 minutes to complete a simple handshake, given
@@
 also have rapidly changing networks at the edges, but the difference is that the
-core of the network is stable traditionally. Where depending on the hour, day,
+core of the network is often stable. Where depending on the hour, day,
 year, the core of the interplanetary network can change drastically. To be able
@@
 except for Earth.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\textwidth]{media/ipn.png}
+  \caption{A sample of the current interplanetary network}
+\end{figure}
 
@@
 confirmations. Other protocols like store and load, and Licklider are used to
-also overcome these problems.
+also overcome these problems. In this paper since no complete implementation of 
+SCPS TP has been created for NS-3, New Reno TCP was used instead.
 
 The rest of the paper remains to simulate these unique challenges, as well as
-see in what conditions do the current protocols work.
+see how protocols operate at these different situations.
 
@@
 
-A significant challenge in creating a IPN is the rapidly changing topology. To
+A significant challenge in modeling a IPN is modeling the rapidly changing topology. To
 simulate these rapid changing conditions a physics engine was created to model
@@
 The first assumption made was that all orbits are circular. Realistically,
-orbits are oval shaped, with the two focal points. A result of this assumption
+orbits are oval shaped, with two focal points. A result of this assumption
 is the elimination of any difference in orbital speed due to Kepler's third law.
+Normally a planet will accelerate or decelerate depending on where it is in 
+its orbit around the sun. 
 The orbital radius was chosen by taking the average orbital radius of the real
@@
 billion meters~\cite{nasa_earth}, and therefore orbits around the sun in a
-circular orbit of radius 149.6 billion meters. Another result of this assumption
+circular orbit of radius 149.6 billion meters in our model. Another result of this assumption
 is that orbits are considered to be centered around the body which it orbits.
@@
 propagation delay in an IPN is much higher than in terrestrial networks, other
-delays such as queuing or processing delays were ignored as being negligible.
+delays such as processing delays were ignored as being negligible.
 
@@
 transmission frequency, giving a transmission wavelength of about 9.99~mm.
-Propagation delay is a simple function of distance and the speed of light. The
-input of the physics engine is a time $t$ from some preset initial conditions
-which outputs the state of all entities.
+Propagation delay is a simple function of distance and the speed of light.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.4\textwidth]{media/sun_orbit.png}
+  \includegraphics[width=0.4\textwidth]{media/moon_orbit.png}
+  \caption{A visual output of the physics model. Left: Mars and Earth orbiting the sun. Right: The ISS (overlapping the center) and the Moon orbiting the Earth}
+\end{figure}
+
+The physics simulation modeled the Sun, Earth, Moon, International Space Station,
+Mars, and the Mars Orbiter. All were able to connect to the network except for the 
+the Sun and the Moon. The input of the physics engine is a time $t$ from some preset initial conditions
+which outputs the state of all entities. Which is fed into the NS-3 simulation.
 
@@
 
-\printbibliography{}
+During the work on this project we contributed to the open source NS-3 project 
+by committing a change to fix a bug which occurred when running the program on 
+different operating systems.
+
+\section{Data}
+The NS-3 simulation ran on the basis of transferring data between Earth and 
+Mars.
+
+Data was scraped from the trace output that was generated while running the 
+ns3 simulation. Trace data was given when a message was queued, dequeued, 
+and received at the destination. This data also included the amount of 
+data being sent in bytes. There is other data on the messages being sent, but 
+was not tracked for the purpose of this project. The trace data was parsed 
+using regex expressions to extract the needed numbers from the string.
+This data was then fed into a pandas dataframe. A unique id had to be 
+generated for each message since once a message was recieved the id was 
+reused. This was done by tracking when each message was received with id 
+$x$ and after incrementing a number by one, say $y$. The unique id was then 
+created by hashing $xy$. Once the data frame was created it had to be merged 
+on itself to have one row of the dataframe represent exactly one message rather 
+then one row be either queue, dequeue, or received.
+
+Time from being queued to received, time being queued, and message time in transit
+were all calculated based off of the data.
+
+\[\text{total\_package\_time} = \text{time\_recieved} - \text{time\_queued}\]
+
+\[\text{time\_queued} = \text{time\_dequeued} - \text{time\_queued}\]
+
+\[\text{total\_time} = \text{time\_received} - \text{time\_dequeued}\]
+
+After the data was processed it was saved in a csv file with one row representing 
+exactly one message that was sent over the network. This file was close to 
+2.5 GB in size.
+
+Statistics were calculated on the entire column of the calculated rows, and 
+bytes sent. Messages were also grouped together which occurred during a similar 
+time window and further averages were calculated to find how the average changed 
+over time of the simulation. These calculations are shown in the graphs in the 
+analysis section.
+
+\section{Analysis}
+
+\subsection{UDP}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.4\textwidth]{media/udp.png}
+  \includegraphics[width=0.4\textwidth]{media/udp_data.png}
+  \caption{UDP analyzed}
+\end{figure}
+
+UDP had an average time from being queued to received of
+195.4 seconds, an average time being queued of
+18.36 seconds, and an average message time in transit
+177.0 seconds. The average amount of data sent was 
+1051.6 bytes. Notice the scientific notation in the 
+corner of the graph. The graph shows a large amount of 
+sensitivity to the motion of the planets which can be 
+seen by the curvature of the graph. At time around 28000
+seconds, the topology suddenly changed. Caused by a satellite
+blocking the path between Mars and Earth. This caused a secondary, 
+less efficient path to be taken. This shows why there are several 
+"layers" of time which it took. They likely represent the different paths 
+which the message took before being sent to its destination. Each quanta in
+the time it took to complete the message 
+represents the smaller jump it takes to get from a satellite to its 
+orbiting entity. The increase in time likely represents that while the 
+direct Earth and Mars link was down the smaller satellites were actually orbiting 
+away from the message destination.
+
+\subsection{TCP}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.4\textwidth]{media/tcp.png}
+  \includegraphics[width=0.4\textwidth]{media/tcp_data.png}
+  \caption{UDP analyzed}
+\end{figure}
+
+\printbibliography
 

### 2024-04-10 draw_orbits.py

@@ -28,7 +28,7 @@ def animate(i):
     t =  2 * SECONDS_IN_DAY * i  
 
     # slow down timer to see moon orbit earth
-    t = 360 * i
+    # t = 360 * i
 
     x_data = []
     y_data = []
@@ -43,8 +43,8 @@ def animate(i):
         y_data.append(datum["y"])
 
     # uncomment to see moon orbit earth
-    ax.set_xlim([data[1]["x"] - EARTH_ORBIT_RAD, data[1]["x"] + EARTH_ORBIT_RAD])
-    ax.set_ylim([data[1]["y"] - EARTH_ORBIT_RAD, data[1]["y"] + EARTH_ORBIT_RAD]) 
+    # ax.set_xlim([data[1]["x"] - EARTH_ORBIT_RAD, data[1]["x"] + EARTH_ORBIT_RAD])
+    # ax.set_ylim([data[1]["y"] - EARTH_ORBIT_RAD, data[1]["y"] + EARTH_ORBIT_RAD]) 
 
     # add it to animation
     scatter.set_offsets(np.column_stack((x_data, y_data)))  
@@ -54,10 +54,10 @@ def animate(i):
 # bootstrap for animated python file
 def animated():
     anim = animation.FuncAnimation(fig, animate, frames = 500, interval = 20, blit = True)
-    # plt.show()
+    plt.show()
 
-    writergif = animation.PillowWriter(fps=30) 
-    anim.save("data/orbit_earth.gif", writer=writergif)
+    # writergif = animation.PillowWriter(fps=30) 
+    # anim.save("data/orbit_earth.gif", writer=writergif)
 
 
 # non animated daily iteration through drawing

### 2024-04-10 paper/paper.tex
added data analysis

@@
 radius of the entity. One of the largest factors contributing to transmission
-error is the distance travelled by the electromagnetic signal between
+error is the distance traveled by the electromagnetic signal between
 transmitting and receiving nodes. By generalizing Friis' transmission
@@
 nodes made it easier to decouple the applications from the network topology and
-allowed us to programatically create arbitrary topologies.
+allowed us to programmatically create arbitrary topologies.
 
@@
 This data was then fed into a pandas dataframe. A unique id had to be 
-generated for each message since once a message was recieved the id was 
+generated for each message since once a message was received the id was 
 reused. This was done by tracking when each message was received with id 
@@
 
-UDP had an average time from being queued to received of
-195.4 seconds, an average time being queued of
-18.36 seconds, and an average message time in transit
-177.0 seconds. The average amount of data sent was 
-1051.6 bytes. Notice the scientific notation in the 
-corner of the graph. The graph shows a large amount of 
+% UDP had an average time from being queued to received of
+% 195.4 seconds, an average time being queued of
+% 18.36 seconds, and an average message time in transit
+% 177.0 seconds.
+
+The average amount of data sent was 
+1051.6 bytes. The number of bytes 
+Notice the scientific notation in the 
+corner of the graph.
+
+% remove this line if stats are re-calculated
+The average time that a message was in transit was close to 216 seconds.
+The time spent waiting in queue was 0.094\% of the time it took for a message to completely 
+transmit.
+
+The graph shows a large amount of 
 sensitivity to the motion of the planets which can be 
@@
 direct Earth and Mars link was down the smaller satellites were actually orbiting 
-away from the message destination.
+away from the message destination. UDP messages spent much more time being queued than 
+the other messages. This may imply that messages were being transferred back and forth 
+between two closely orbiting objects, which could explain the many different layers 
+which it can take for a message to be sent. It could also be that since the average 
+message is larger than TCP or New Reno it takes longer for the message to be sent 
+and it is transmitting.
 
@@
   \includegraphics[width=0.4\textwidth]{media/tcp_data.png}
-  \caption{UDP analyzed}
+  \caption{TCP analyzed}
 \end{figure}
+
+The average amount of bytes transferred is 394 bytes. This is likely from 
+the difference in the amounts of bytes transferred in an ACK message compared 
+to when data is being sent.
+
+Unlike in UDP the motion of the planets is not observable in the graph. However this is 
+because there are several different paths which the TCP message could have taken which can be seen. 
+Since TCP requires a higher amount of base time to establish a connection compared to UDP, this 
+means that when it "hops" between entities it adds a larger base time.
+
+At the start of the time we can see 4 separate paths. This would be the different times that it would take 
+for any of the 4 paths between Earth and Mars, that is Earth to the ISS to the Mars Orbiter then to Mars,
+Earth to the Mars Orbiter to Mars, Earth to the ISS to Mars, and finally Earth to Mars. After one of the 
+entities is blocking a direct signal it becomes only that 3 of these options possible, with likely only two actual 
+paths being viable. With less available 
+bandwidth it becomes slower on average. Smaller time may also be correlated to the minority of messages which were
+sent with a smaller message size.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.4\textwidth]{media/tcp_future.png}
+  \includegraphics[width=0.4\textwidth]{media/tcp_future_data.png}
+  \caption{TCP in the future analyzed}
+\end{figure}
+
+The second set of graphs show TCP in some time in the future. We can see that the time that messages
+take to transfer does in fact change over time as the entities change distances. We can also see a similar
+pattern of behaviour changing as some topology changes in the model.
+
+\subsection{New Reno}
+
+New Reno, a modified version of TCP used to help elevate congestion ~\cite{Nahar2016}, was used 
+in place of SCP TP. New Reno behaved very similar to TCP. Its average data 
+transfer was 407.5 bytes.
+
+It preformed marginally better in its time from being queued to the message received. This 
+was mostly due to a decrease in time that it was in queue, which is to be expected. There was 
+a very marginal increase in time in transit however.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.4\textwidth]{media/new_reno.png}
+  \includegraphics[width=0.4\textwidth]{media/new_reno_data.png}
+  \caption{TCP in the future analyzed}
+\end{figure}
+
+Similar to TCP we can see the two stages which occur in the simulation. One where most of 
+the messages were likely being routed from directly from Earth to Mars, which then gets transformed 
+into having to route through one of the external satellites orbiting each entity, which decreases the 
+average time it takes to transmit the message.
+
+\subsection{Protocol Comparison}
+
+Since a large amount of propagation delay dominates the amount of time that it takes for messages total 
+transmit time from being queued to being received at the appropriate node, other increases in performance 
+mean only a slight decrease in time sending the message. An important statistic to pay attention to is that 
+UDP almost always sent a message with a size of 1052 bytes. This means that even though TCP, New Reno, and 
+UDP had sent a similar number of messages, the actual throughput for UDP was significantly higher, which 
+is much more significant in an IPN situation. In the data graphs we can see that not only was the maximum 
+data being sent was less than UDP, but also had to send messages of very little size, most likely acknowledgements,
+as well. New Reno was seen to improve on the queuing time compared to both TCP and UDP. Which would mean in a 
+high reliability situation and in situations where higher traffic is sent through a network with a relatively low 
+bandwidth, like a IPN, where satellites are limited by how much data they can transmit, New Reno's effect can be 
+more significant. This should only be used in a connection where propagation delay is less of an issue like 
+between planet and satellite communication. Queuing time in UDP took a more significant percentage of its 
+time than both TCP and New Reno. In general however, UDP is the best option for primary communication between 
+Mars and Earth communication.
+
+\section{Future Work}
+
+Future work that could be done would be to implement or partially implement SCP TP as a network 
+protocol which could be used and compared to the protocols analyzed in this paper. Other protocols 
+like the Linklider or load and store protocols could also be implemented, to create a more realistic 
+IPN.
+
+The physics engine could be expanded to included 3D orbits, particularly around planets, where networks of 
+satellites could be created to create a proper level of communication that is maintained today, with geocentric 
+and polar orbiting satellites. Connection points from a planet to off the planet could have a direction which it 
+needs to send a message in for it to be able to connect. For example a rover on Mars in our simulation could technically 
+send a message in any direction, even if they were sending it through the surface of the planet.The error model in 
+the physics engine could be expanded to include interference like solar radiation, atmosphere, and other interference.
+
+Other stats could be extracted from the NS-3 model such as packet loss, message path, handshake time. It could expand 
+so that the network was not just focused on Earth to Mars communication. It could be expanded to include other celestial 
+objects in the solar system, as well as potential future objects.
+
+The NS-3 simulation could dynamically choose the network protocol to use based on the level of error, message size,
+estimated propagation error, and other factors. This could build a model which uses not only one protocol but all of them 
+and leverages the advantages of each when it is necessary. This combined with a more sophisticated model of the solar 
+system would create a very competent simulation of the current operation of the IPN, and could then be expanded or tweaked 
+to find if expansions to it could be handled properly. This would need a combination of all of the suggestions above 
+to be successful.
+
+\section{Conclusion}
+
+In this paper we outlined how we created a model of an interplanetary internet (IPN) using a physics engine which fed into 
+a NS-3 simulation. The physics simulation used a simplified model of orbiting planets and satellites as well as physics calculations 
+to create data which would be the output for any input time $t$. The NS-3 model used a point to point base which programmatically
+took the data from the physics engine and created nodes which acted as routers. A simulation was ran which transfered data between 
+Earth and Mars, using satellites and other entities to relay messages around blocked connections. Analysis was done on three
+different network protocols used in the IPN, that is UDP, TCP and New Reno. 
+UDP was found to have the most throughput out of all of the three protocols, while New Reno reduced the amount of time a message 
+was queued.
 

### 2024-04-10 paper/paper.tex
changed title
remove appendix reference
update the paper and add missing stats

@@
 
-\title{Analyzing Network Transfer Protocols on a Simulated Interplanetary Internet}
+\title{Simulating an Interplanetary Internet}
 \author{Cody Auch, Mark Lysack, Jacob Janzen}
@@
 
-\newcommand{\CC}{C\nolinebreak\hspace{-.05em}\raisebox{.4ex}{\tiny\bf +}\nolinebreak\hspace{-.10em}\raisebox{.4ex}{\tiny\bf +}}
+\newcommand{\CC}{C\nolinebreak\hspace{-.05em}\raisebox{.4ex}{\tiny\bf
+    +}\nolinebreak\hspace{-.10em}\raisebox{.4ex}{\tiny\bf +}}
 \def\CC{{C\nolinebreak[4]\hspace{-.05em}\raisebox{.4ex}{\tiny\bf ++}}}
@@
   from them that make them a unique challenge compared to shorter-range computer
-  networks, and how our current network protocols can be applied in this application. In
-  addition, we look at how one can use conventional network simulators to test
-  interplanetary network implementations. We used the network simulator
-  NS-3 ~\cite{ns-3} to test various transport layer protocols against each other
-  in a network topology that we designed in the network simulator to emulate
-  long-distance communications. Using this network simulation, we were able to
-  compare UDP, TCP, and TCP New Reno protocols against each other to see how
-  they each performed across such large distances with high delays. UDP preformed the 
-  best with New Reno and TCP closely paired together.
+  networks, and how our current network protocols can be applied in this
+  application. In addition, we look at how one can use conventional network
+  simulators to test interplanetary network implementations. We used the network
+  simulator NS-3 ~\cite{ns-3} to test various transport layer protocols against
+  each other in a network topology that we designed in the network simulator to
+  emulate long-distance communications. Using this network simulation, we were
+  able to compare UDP, TCP, and TCP New Reno protocols against each other to see
+  how they each performed across such large distances with high delays. UDP
+  preformed the best with New Reno and TCP closely paired together.
 \end{abstract}
@@
 An interplanetary network is simply a network which where the nodes are
-distributed across the surface or in the orbits of different planets or other celestial objects.
-In the year of
-2024 the current interplanetary is limited to Mars, Earth, Earth orbit, Moon orbit, and
-satellites orbiting the sun. The interplanetary network is also expanded to deep
-space satellites, like the Voyager probes. A sample message path across the
-network could be that a message is passed from a large network antenna on earth
-to a earth orbiting satellite, then when it has direct line of
-sight with MAVEN (a satellite orbiting mars) it sends the message. MAVEN then relays the message to a
-Mars rover like Odyssey. This description hides the complexity which arises in
-an interplanetary network which may seem like a extremely simple network.
+distributed across the surface or in the orbits of different planets or other
+celestial objects. In the year of 2024 the current interplanetary is limited to
+Mars, Earth, Earth orbit, Moon orbit, and satellites orbiting the sun. The
+interplanetary network is also expanded to deep space satellites, like the
+Voyager probes. A sample message path across the network could be that a message
+is passed from a large network antenna on earth to a earth orbiting satellite,
+then when it has direct line of sight with MAVEN (a satellite orbiting mars) it
+sends the message. MAVEN then relays the message to a Mars rover like Odyssey.
+This description hides the complexity which arises in an interplanetary network
+which may seem like a extremely simple network.
 
-In a traditional network a majority of the network traffic will occur over a wired
-connection, especially for transfers of large amounts of data.
-This means that the error is relatively low. This is no longer the
-case in the case of satellites sending messages over potentially millions of
-kilometers. Atmosphere in the Earth orbit, solar radiation, and dust cause
-increased error in sending the message between entities. Inverse square law
-dictates that over these large periods the energy needed to send the message so
-that it can be read at its destination. Even the Doppler effect is needed to be
-accounted for to tune into radio frequencies on the ISS. ~\cite{qslCompensatingDoppler} High error means that
-sending a message can no longer be assumed to be sent reliably and has to be
-compensated for. Beyond high error, large distances between entities and the
+In a traditional network a majority of the network traffic will occur over a
+wired connection, especially for transfers of large amounts of data. This means
+that the error is relatively low. This is no longer the case in the case of
+satellites sending messages over potentially millions of kilometers. Atmosphere
+in the Earth orbit, solar radiation, and dust cause increased error in sending
+the message between entities. Inverse square law dictates that over these large
+periods the energy needed to send the message so that it can be read at its
+destination. Even the Doppler effect is needed to be accounted for to tune into
+radio frequencies on the ISS.~\cite{qslCompensatingDoppler} High error means
+that sending a message can no longer be assumed to be sent reliably and has to
+be compensated for. Beyond high error, large distances between entities and the
 physical limitation that messages cannot travel faster than the speed of light.
-Light, or any other energy takes approximately 8 minutes to get from the Sun to Earth, ~\cite{nasa_earth} and
-thus for a round trip will take 16 minutes to complete a simple handshake, given
-that the messages were sent correctly.
+Light, or any other energy takes approximately 8 minutes to get from the Sun to
+Earth,~\cite{nasa_earth} and thus for a round trip will take 16 minutes to
+complete a simple handshake, given that the messages were sent correctly.
 
@@
 also have rapidly changing networks at the edges, but the difference is that the
-core of the network is often stable. Where depending on the hour, day,
-year, the core of the interplanetary network can change drastically. To be able
-to connect to the network there are no well known places to access the network,
-except for Earth.
+core of the network is often stable. Where depending on the hour, day, year, the
+core of the interplanetary network can change drastically. To be able to connect
+to the network there are no well known places to access the network, except for
+Earth.
 
@@
 
-A significant challenge in modeling a IPN is modeling the rapidly changing topology. To
-simulate these rapid changing conditions a physics engine was created to model
-the orbits of several entities in the solar system. Several assumptions were
-made in the simulator for the sake of simplicity as the goal of the simulator is
-not to create a perfect model of the solar system and the entities inside of the
-solar system, but rather conditions which a IPN would exist in.
+A significant challenge in modeling a IPN is modeling the rapidly changing
+topology. To simulate these rapid changing conditions a physics engine was
+created to model the orbits of several entities in the solar system. Several
+assumptions were made in the simulator for the sake of simplicity as the goal of
+the simulator is not to create a perfect model of the solar system and the
+entities inside of the solar system, but rather conditions which a IPN would
+exist in.
 
 The first assumption made was that all orbits are circular. Realistically,
-orbits are oval shaped, with two focal points. A result of this assumption
-is the elimination of any difference in orbital speed due to Kepler's third law.
-Normally a planet will accelerate or decelerate depending on where it is in 
-its orbit around the sun. 
-The orbital radius was chosen by taking the average orbital radius of the real
-life celestial bodies. For example Earth's average orbital radius is 149.6
-billion meters~\cite{nasa_earth}, and therefore orbits around the sun in a
-circular orbit of radius 149.6 billion meters in our model. Another result of this assumption
-is that orbits are considered to be centered around the body which it orbits.
-The Mars orbit has a focal point that is further in space, making it more oval
-shaped than the earth orbit.~\cite{nasa_mars} While this could make changes to a
-simulated network in some very specific situations, the primary behaviour of a
-solar system is maintained, like retrograde orbits, periodical alignment,
-predictable behaviour, and recursive orbits of sub orbiting object.
+orbits are oval shaped, with two focal points. A result of this assumption is
+the elimination of any difference in orbital speed due to Kepler's third law.
+Normally a planet will accelerate or decelerate depending on where it is in its
+orbit around the sun. The orbital radius was chosen by taking the average
+orbital radius of the real life celestial bodies. For example Earth's average
+orbital radius is 149.6 billion meters~\cite{nasa_earth}, and therefore orbits
+around the sun in a circular orbit of radius 149.6 billion meters in our model.
+Another result of this assumption is that orbits are considered to be centered
+around the body which it orbits. The Mars orbit has a focal point that is
+further in space, making it more oval shaped than the earth
+orbit.~\cite{nasa_mars} While this could make changes to a simulated network in
+some very specific situations, the primary behaviour of a solar system is
+maintained, like retrograde orbits, periodical alignment, predictable behaviour,
+and recursive orbits of sub orbiting object.
 
@@
 message, and to calculate the message propagation delay for the message to be
-sent. Line of sight was achieved using simple geometry calculations. (See
-appendix) All entities are able to block signals if the ray passes through the
-radius of the entity. One of the largest factors contributing to transmission
-error is the distance traveled by the electromagnetic signal between
-transmitting and receiving nodes. By generalizing Friis' transmission
-formula\cite{Friis} for isotropic emitters and receivers, we can obtain the
-ratio of signal intensity received to signal intensity transmitted, known as
-Free Space Path Loss (FSPL). We can use FSPL as our transmission error rate,
-giving the following expression: $Error Rate = {(\frac{4 \pi d}{\lambda})}^2$
-where $d$ is the distance between nodes in metres and $\lambda$ is the
-wavelength of the electromagnetic wave used in transmission. As much of space
-communications utilizes the Ka-band of frequencies (27--40
-GHz)\cite{Morabito_Hastrup} we opted to use a transmission frequency of 30Ghz.
-The transmission wavelength can be then calculated according to
-$\lambda = \frac{c}{f}$, where $c$ is the speed of light and $f$ is the
+sent. Line of sight was achieved using simple geometry calculations. All
+entities are able to block signals if the ray passes through the radius of the
+entity. One of the largest factors contributing to transmission error is the
+distance traveled by the electromagnetic signal between transmitting and
+receiving nodes. By generalizing Friis' transmission formula\cite{Friis} for
+isotropic emitters and receivers, we can obtain the ratio of signal intensity
+received to signal intensity transmitted, known as Free Space Path Loss (FSPL).
+We can use FSPL as our transmission error rate, giving the following expression:
+$Error Rate = {(\frac{4 \pi d}{\lambda})}^2$ where $d$ is the distance between
+nodes in metres and $\lambda$ is the wavelength of the electromagnetic wave used
+in transmission. As much of space communications utilizes the Ka-band of
+frequencies (27--40 GHz)\cite{Morabito_Hastrup} we opted to use a transmission
+frequency of 30Ghz. The transmission wavelength can be then calculated according
+to $\lambda = \frac{c}{f}$, where $c$ is the speed of light and $f$ is the
 transmission frequency, giving a transmission wavelength of about 9.99~mm.
@@
   \includegraphics[width=0.4\textwidth]{media/moon_orbit.png}
-  \caption{A visual output of the physics model. Left: Mars and Earth orbiting the sun. Right: The ISS (overlapping the center) and the Moon orbiting the Earth}
+  \caption{A visual output of the physics model. Left: Mars and Earth orbiting
+    the sun. Right: The ISS (overlapping the center) and the Moon orbiting the
+    Earth}
 \end{figure}
 
-The physics simulation modeled the Sun, Earth, Moon, International Space Station,
-Mars, and the Mars Orbiter. All were able to connect to the network except for the 
-the Sun and the Moon. The input of the physics engine is a time $t$ from some preset initial conditions
-which outputs the state of all entities. Which is fed into the NS-3 simulation.
+The physics simulation modeled the Sun, Earth, Moon, International Space
+Station, Mars, and the Mars Orbiter. All were able to connect to the network
+except for the the Sun and the Moon. The input of the physics engine is a time
+$t$ from some preset initial conditions which outputs the state of all entities.
+Which is fed into the NS-3 simulation.
 
@@
 
-Time from being queued to received, time being queued, and message time in transit
-were all calculated based off of the data.
+Time from being queued to received, time being queued, and message time in
+transit were all calculated based off of the data.
 
@@
 
-After the data was processed it was saved in a csv file with one row representing 
-exactly one message that was sent over the network. This file was close to 
-2.5 GB in size.
+After the data was processed it was saved in a csv file with one row
+representing exactly one message that was sent over the network. This file was
+close to 2.5 GB in size.
 
@@
 
-% UDP had an average time from being queued to received of
-% 195.4 seconds, an average time being queued of
-% 18.36 seconds, and an average message time in transit
-% 177.0 seconds.
+UDP had an average time from being queued to received of 283.64 seconds, an
+average time being queued of 21.92 seconds, and an average message time in
+transit 261.72 seconds. The average amount of data sent was 1052.0 bytes. Notice
+the scientific notation in the corner of the graph.
 
-The average amount of data sent was 
-1051.6 bytes. The number of bytes 
-Notice the scientific notation in the 
-corner of the graph.
-
-% remove this line if stats are re-calculated
-The average time that a message was in transit was close to 216 seconds.
-The time spent waiting in queue was 0.094\% of the time it took for a message to completely 
-transmit.
-
-The graph shows a large amount of 
-sensitivity to the motion of the planets which can be 
-seen by the curvature of the graph. At time around 28000
-seconds, the topology suddenly changed. Caused by a satellite
-blocking the path between Mars and Earth. This caused a secondary, 
-less efficient path to be taken. This shows why there are several 
-"layers" of time which it took. They likely represent the different paths 
-which the message took before being sent to its destination. Each quanta in
-the time it took to complete the message 
-represents the smaller jump it takes to get from a satellite to its 
-orbiting entity. The increase in time likely represents that while the 
-direct Earth and Mars link was down the smaller satellites were actually orbiting 
-away from the message destination. UDP messages spent much more time being queued than 
-the other messages. This may imply that messages were being transferred back and forth 
-between two closely orbiting objects, which could explain the many different layers 
-which it can take for a message to be sent. It could also be that since the average 
-message is larger than TCP or New Reno it takes longer for the message to be sent 
-and it is transmitting.
+The graph shows a large amount of sensitivity to the motion of the planets which
+can be seen by the curvature of the graph. At time around 28000 seconds, the
+topology suddenly changed. Caused by a satellite blocking the path between Mars
+and Earth. This caused a secondary, less efficient path to be taken. This shows
+why there are several ``layers'' of time which it took. They likely represent the
+different paths which the message took before being sent to its destination.
+Each quanta in the time it took to complete the message represents the smaller
+jump it takes to get from a satellite to its orbiting entity. The increase in
+time likely represents that while the direct Earth and Mars link was down the
+smaller satellites were actually orbiting away from the message destination. UDP
+messages spent much more time being queued than the other messages. This may
+imply that messages were being transferred back and forth between two closely
+orbiting objects, which could explain the many different layers which it can
+take for a message to be sent. It could also be that since the average message
+is larger than TCP or New Reno it takes longer for the message to be sent and it
+is transmitting.
 
@@
 
-The average amount of bytes transferred is 394 bytes. This is likely from 
-the difference in the amounts of bytes transferred in an ACK message compared 
-to when data is being sent.
+The average amount of bytes transferred is 320.44 bytes. This is likely from the
+difference in the amounts of bytes transferred in an ACK message compared to
+when data is being sent.
 
-Unlike in UDP the motion of the planets is not observable in the graph. However this is 
-because there are several different paths which the TCP message could have taken which can be seen. 
-Since TCP requires a higher amount of base time to establish a connection compared to UDP, this 
-means that when it "hops" between entities it adds a larger base time.
+Unlike in UDP the motion of the planets is not observable in the graph. However
+this is because there are several different paths which the TCP message could
+have taken which can be seen. Since TCP requires a higher amount of base time to
+establish a connection compared to UDP, this means that when it ``hops'' between
+entities it adds a larger base time.
 
-At the start of the time we can see 4 separate paths. This would be the different times that it would take 
-for any of the 4 paths between Earth and Mars, that is Earth to the ISS to the Mars Orbiter then to Mars,
-Earth to the Mars Orbiter to Mars, Earth to the ISS to Mars, and finally Earth to Mars. After one of the 
-entities is blocking a direct signal it becomes only that 3 of these options possible, with likely only two actual 
-paths being viable. With less available 
-bandwidth it becomes slower on average. Smaller time may also be correlated to the minority of messages which were
-sent with a smaller message size.
+At the start of the time we can see 4 separate paths. This would be the
+different times that it would take for any of the 4 paths between Earth and
+Mars, that is Earth to the ISS to the Mars Orbiter then to Mars, Earth to the
+Mars Orbiter to Mars, Earth to the ISS to Mars, and finally Earth to Mars. After
+one of the entities is blocking a direct signal it becomes only that 3 of these
+options possible, with likely only two actual paths being viable. With less
+available bandwidth it becomes slower on average. Smaller time may also be
+correlated to the minority of messages which were sent with a smaller message
+size.
 
@@
 
-The second set of graphs show TCP in some time in the future. We can see that the time that messages
-take to transfer does in fact change over time as the entities change distances. We can also see a similar
-pattern of behaviour changing as some topology changes in the model.
+The second set of graphs show TCP in some time in the future. We can see that
+the time that messages take to transfer does in fact change over time as the
+entities change distances. We can also see a similar pattern of behaviour
+changing as some topology changes in the model.
 
@@
 
-New Reno, a modified version of TCP used to help elevate congestion ~\cite{Nahar2016}, was used 
-in place of SCP TP. New Reno behaved very similar to TCP. Its average data 
-transfer was 407.5 bytes.
+New Reno, a modified version of TCP used to help reduce
+congestion~\cite{Nahar2016}, was used in place of SCP TP. New Reno behaved very
+similar to TCP. Its average data transfer was 307.18 bytes.
 
-It preformed marginally better in its time from being queued to the message received. This 
-was mostly due to a decrease in time that it was in queue, which is to be expected. There was 
-a very marginal increase in time in transit however.
+It preformed marginally better in its time from being queued to the message
+received. This was mostly due to a decrease in time that it was in queue, which
+is to be expected. There was a very marginal increase in time in transit
+however.
 
@@
 
-Similar to TCP we can see the two stages which occur in the simulation. One where most of 
-the messages were likely being routed from directly from Earth to Mars, which then gets transformed 
-into having to route through one of the external satellites orbiting each entity, which decreases the 
-average time it takes to transmit the message.
+Similar to TCP we can see the two stages which occur in the simulation. One
+where most of the messages were likely being routed from directly from Earth to
+Mars, which then gets transformed into having to route through one of the
+external satellites orbiting each entity, which decreases the average time it
+takes to transmit the message.
 
@@
 
-Since a large amount of propagation delay dominates the amount of time that it takes for messages total 
-transmit time from being queued to being received at the appropriate node, other increases in performance 
-mean only a slight decrease in time sending the message. An important statistic to pay attention to is that 
-UDP almost always sent a message with a size of 1052 bytes. This means that even though TCP, New Reno, and 
-UDP had sent a similar number of messages, the actual throughput for UDP was significantly higher, which 
-is much more significant in an IPN situation. In the data graphs we can see that not only was the maximum 
-data being sent was less than UDP, but also had to send messages of very little size, most likely acknowledgements,
-as well. New Reno was seen to improve on the queuing time compared to both TCP and UDP. Which would mean in a 
-high reliability situation and in situations where higher traffic is sent through a network with a relatively low 
-bandwidth, like a IPN, where satellites are limited by how much data they can transmit, New Reno's effect can be 
-more significant. This should only be used in a connection where propagation delay is less of an issue like 
-between planet and satellite communication. Queuing time in UDP took a more significant percentage of its 
-time than both TCP and New Reno. In general however, UDP is the best option for primary communication between 
-Mars and Earth communication.
+Since a large amount of propagation delay dominates the amount of time that it
+takes for messages total transmit time from being queued to being received at
+the appropriate node, other increases in performance mean only a slight decrease
+in time sending the message. An important statistic to pay attention to is that
+UDP almost always sent a message with a size of 1052 bytes. This means that even
+though TCP, New Reno, and UDP had sent a similar number of messages, the actual
+throughput for UDP was significantly higher, which is much more significant in
+an IPN situation. In the data graphs we can see that not only was the maximum
+data being sent was less than UDP, but also had to send messages of very little
+size, most likely acknowledgments, as well. New Reno was seen to improve on the
+queuing time compared to both TCP and UDP. Which would mean in a high
+reliability situation and in situations where higher traffic is sent through a
+network with a relatively low bandwidth, like a IPN, where satellites are
+limited by how much data they can transmit, New Reno's effect can be more
+significant. This should only be used in a connection where propagation delay is
+less of an issue like between planet and satellite communication. Queuing time
+in UDP took a more significant percentage of its time than both TCP and New
+Reno. In general however, UDP is the best option for primary communication
+between Mars and Earth communication.
 
@@
 
-Future work that could be done would be to implement or partially implement SCP TP as a network 
-protocol which could be used and compared to the protocols analyzed in this paper. Other protocols 
-like the Linklider or load and store protocols could also be implemented, to create a more realistic 
-IPN.
+Future work that could be done would be to implement or partially implement SCP
+TP as a network protocol which could be used and compared to the protocols
+analyzed in this paper. Other protocols like the Linklider or load and store
+protocols could also be implemented, to create a more realistic IPN.
 
-The physics engine could be expanded to included 3D orbits, particularly around planets, where networks of 
-satellites could be created to create a proper level of communication that is maintained today, with geocentric 
-and polar orbiting satellites. Connection points from a planet to off the planet could have a direction which it 
-needs to send a message in for it to be able to connect. For example a rover on Mars in our simulation could technically 
-send a message in any direction, even if they were sending it through the surface of the planet.The error model in 
-the physics engine could be expanded to include interference like solar radiation, atmosphere, and other interference.
+The physics engine could be expanded to included 3-D orbits, particularly around
+planets, where networks of satellites could be created to create a proper level
+of communication that is maintained today, with geocentric and polar orbiting
+satellites. Connection points from a planet to off the planet could have a
+direction which it needs to send a message in for it to be able to connect. For
+example a rover on Mars in our simulation could technically send a message in
+any direction, even if they were sending it through the surface of the
+planet.The error model in the physics engine could be expanded to include
+interference like solar radiation, atmosphere, and other interference.
 
-Other stats could be extracted from the NS-3 model such as packet loss, message path, handshake time. It could expand 
-so that the network was not just focused on Earth to Mars communication. It could be expanded to include other celestial 
+Other stats could be extracted from the NS-3 model such as packet loss, message
+path, handshake time. It could expand so that the network was not just focused
+on Earth to Mars communication. It could be expanded to include other celestial
 objects in the solar system, as well as potential future objects.
 
-The NS-3 simulation could dynamically choose the network protocol to use based on the level of error, message size,
-estimated propagation error, and other factors. This could build a model which uses not only one protocol but all of them 
-and leverages the advantages of each when it is necessary. This combined with a more sophisticated model of the solar 
-system would create a very competent simulation of the current operation of the IPN, and could then be expanded or tweaked 
-to find if expansions to it could be handled properly. This would need a combination of all of the suggestions above 
-to be successful.
+The NS-3 simulation could dynamically choose the network protocol to use based
+on the level of error, message size, estimated propagation error, and other
+factors. This could build a model which uses not only one protocol but all of
+them and leverages the advantages of each when it is necessary. This combined
+with a more sophisticated model of the solar system would create a very
+competent simulation of the current operation of the IPN, and could then be
+expanded or tweaked to find if expansions to it could be handled properly. This
+would need a combination of all of the suggestions above to be successful.
 
@@
 
-In this paper we outlined how we created a model of an interplanetary internet (IPN) using a physics engine which fed into 
-a NS-3 simulation. The physics simulation used a simplified model of orbiting planets and satellites as well as physics calculations 
-to create data which would be the output for any input time $t$. The NS-3 model used a point to point base which programmatically
-took the data from the physics engine and created nodes which acted as routers. A simulation was ran which transfered data between 
-Earth and Mars, using satellites and other entities to relay messages around blocked connections. Analysis was done on three
-different network protocols used in the IPN, that is UDP, TCP and New Reno. 
-UDP was found to have the most throughput out of all of the three protocols, while New Reno reduced the amount of time a message 
-was queued.
+In this paper we outlined how we created a model of an interplanetary internet
+(IPN) using a physics engine which fed into a NS-3 simulation. The physics
+simulation used a simplified model of orbiting planets and satellites as well as
+physics calculations to create data which would be the output for any input time
+$t$. The NS-3 model used a point to point base which programmatically took the
+data from the physics engine and created nodes which acted as routers. A
+simulation was ran which transfered data between Earth and Mars, using
+satellites and other entities to relay messages around blocked connections.
+Analysis was done on three different network protocols used in the IPN, that is
+UDP, TCP and New Reno. UDP was found to have the most throughput out of all of
+the three protocols, while New Reno reduced the amount of time a message was
+queued.
 
-\printbibliography
+\printbibliography{}
 

### 2024-04-10 process_data.py

@@ -1,5 +1,6 @@
 import math
 import re
+import sys
 import pandas as pd
 import numpy as np
 import matplotlib.pyplot as plt
@@ -7,11 +8,12 @@ import matplotlib.pyplot as plt
 
 # take trace file and turn it into a dataframe
 def process_trace():
-    file1 = open('network-sim.tr', 'r')
+    file1 = open(sys.argv[1], 'r')
     Lines = file1.readlines()
 
     processed_data = []
 
+    ids = {}
     for line in Lines:
         try:
             TYPE_INDEX = 1
@@ -19,7 +21,7 @@ def process_trace():
             NODE_INDEX = 3
             INTERFACE_INDEX = 4
 
-            res = re.compile("^([r\\+-]) ([0-9]*\.[0-9]*|[0-9]*[0-9]*) \/NodeList\/([0-9]*)\/DeviceList\/([0-9]*)").search(line)
+            res = re.compile("^([r\\+-]) ([0-9]*\\.?[0-9]*) \\/NodeList\\/([0-9]*)\\/DeviceList\\/([0-9]*)").search(line)
 
             type = res.group(TYPE_INDEX)
             time = res.group(TIME_INDEX)
@@ -30,7 +32,12 @@ def process_trace():
 
             bytes = re.compile("(?: length: )(.*?(?= ))").search(line).group(1)
 
-            processed_data.append([type, time, node, interface, bytes, id])
+            if id not in ids and type == '+':
+                ids[id] = 0
+            elif id in ids and type == '+':
+                ids[id] += 1
+
+            processed_data.append([type, time, node, interface, bytes, hash(f"{id}t{ids[id]}")])
 
         except Exception as e:
             print("exception encountered")
@@ -83,18 +90,23 @@ def expand_dataframe(df):
     df[["time_queued", "time_dequeued", "time_received"]] = df[["time_queued", "time_dequeued", "time_received"]].astype(float)
 
     # calculate time from being queued to received
-    df["total_package_time"] = df["time_received"] - df["time_queued"]
+    df["total_package_time"] = (df["time_received"] - df["time_queued"]) * 26
 
     # calculate time being queued
-    df["time_queued"] = df["time_dequeued"] - df["time_queued"]
+    df["time_queued"] = (df["time_dequeued"] - df["time_queued"]) * 26
 
     # calculate the message time in transit
-    df["total_time"] = df["time_received"] -  df["time_dequeued"]
[…]

### 2024-04-10 process_data.py
update readme

@@ -184,5 +184,5 @@ def save_dataframe():
 
 # for testing
 save_dataframe()
-# get_graphs()
+get_graphs()
 get_statistics()

### 2024-04-10 paper/paper.tex
fix typo
fix some details

@@
 \title{Simulating an Interplanetary Internet}
-\author{Cody Auch, Mark Lysack, Jacob Janzen}
+\author{Cody Auch, Jacob Janzen, Mark Lysack}
 \date{\today}
@@
   simulators to test interplanetary network implementations. We used the network
-  simulator NS-3 ~\cite{ns-3} to test various transport layer protocols against
+  simulator NS-3~\cite{ns-3} to test various transport layer protocols against
   each other in a network topology that we designed in the network simulator to
@@
   preformed the best with New Reno and TCP closely paired together.
+
+  Our implementation can be found
+  at \url{https://github.com/codyauch/IPN-Project}. The physics simulation was
+  mostly designed by Mark and Cody. The network simulation in NS-3 was done by
+  Jacob, and the analysis was done by Mark.
 \end{abstract}
@@
 orbits are oval shaped, with two focal points. A result of this assumption is
-the elimination of any difference in orbital speed due to Kepler's third law.
+the elimination of any difference in orbital speed due to Kepler's second law.
 Normally a planet will accelerate or decelerate depending on where it is in its
@@
 
-The graph shows a large amount of sensitivity to the motion of the planets which
-can be seen by the curvature of the graph. At time around 28000 seconds, the
-topology suddenly changed. Caused by a satellite blocking the path between Mars
-and Earth. This caused a secondary, less efficient path to be taken. This shows
-why there are several ``layers'' of time which it took. They likely represent the
-different paths which the message took before being sent to its destination.
-Each quanta in the time it took to complete the message represents the smaller
-jump it takes to get from a satellite to its orbiting entity. The increase in
-time likely represents that while the direct Earth and Mars link was down the
-smaller satellites were actually orbiting away from the message destination. UDP
-messages spent much more time being queued than the other messages. This may
-imply that messages were being transferred back and forth between two closely
-orbiting objects, which could explain the many different layers which it can
-take for a message to be sent. It could also be that since the average message
-is larger than TCP or New Reno it takes longer for the message to be sent and it
-is transmitting.
+We believe that UDP sends a large number of messages at the start that are all queued, processed, and sent.
+After those messages are sent, the next batch is prepared immediately.
+This curve could be caused by the queuing delay from the large number of messages that are sent.
+UDP does not wait for messages to be received and will therefore create a large backlog which slows everything down when dequeuing messages.
 
@@
 data from the physics engine and created nodes which acted as routers. A
-simulation was ran which transfered data between Earth and Mars, using
+simulation was ran which transferred data between Earth and Mars, using
 satellites and other entities to relay messages around blocked connections.

### 2024-04-10 paper/paper.tex
added udp stuff
add UDP analysis

@@
 
-We believe that UDP sends a large number of messages at the start that are all queued, processed, and sent.
-After those messages are sent, the next batch is prepared immediately.
-This curve could be caused by the queuing delay from the large number of messages that are sent.
-UDP does not wait for messages to be received and will therefore create a large backlog which slows everything down when dequeuing messages.
+The message size is the same across all of the messages. This makes sense as the message size do not need to be different across different messages. This is different compared to TCP and New Reno.
+
+On the left side of the graph before approximately $t = 25000$ this is where
+the planet most likely has a direct path between Earth and Mars. This has a
+steady pattern of a batch of messages are queued, they are sent, and after they
+are sent the next batch of messages are queued, and calculated the propagation
+delay at the time that it is queued. This leads to the clusters which can be seen.
+The curvature of the delay is most likely attributed to the movement of Earth
+relative to Mars, that is that they are getting closer together. Queueing delay is likely small in this section suggested by the tight groupings.
+
+At approximately time $t = 25000$ the direct path between Earth and Mars is
+blocked. This means that the messages have to be routed through a secondary
+satellite. The vertical gap between the horizontal lines is roughly 0.001
+seconds. This is the same time as it takes for a message to take for a message
+to send between Earth and the ISS and similarly between Mars and the Mars
+Orbiter. The reason for the multiple horizontal lines is that messages may be
+passed several times between the planet and its satellite before it is finally
+delivered across the large propagation delay between the planets Earth and Mars.
+The propagation delay could be seen through the length of the horizontal lines.
+This is representative of the time it takes to clear each batch when it is sent.
+Once an entire batch is sent it will re-calculate the propagation delay which
+will then be represented in new different horizontal lines. The long lines show
+that there was a large spike in queuing delay as the topology changed which
+became smaller as it cleared. The curvature of this curve likely represents the
+change in the relative distance between the satellites orbiting Earth and Mars
+becoming further apart. This is clear here due to the extremely small variation
+in propagation delay.
+
+This analysis cannot be 100\% verified due to the data mined from NS-3 did not give a ton of information about the messages which were sent on our simulated network.
 

### 2024-04-10 paper/paper.tex
idk

@@
 
-In the past decade interest in returning back to space has increased
+In the past decade, interest in returning back to space has increased
 internationally. Many organizations have started sending remote vehicles into
@@
 In their 2017 Masters Thesis, Md Monjurul Islam Khan~\cite{Khan2017} from
-University of Manitoba wrote on how to create a satellite network simulation
-based on NS-3. They used a point to point NS-3 module to create a network of 2
-levels of orbiting satellites and fixed points on the earth which acts as
-uplinks to the network. The paper analyzed TCP HighSpeed against TCP Vegas slow
-start. However this paper stops at expanding the network beyond earth orbit, and
-rather focuses on Earth orbit communication.
+University of Manitoba wrote about how one can create a satellite network
+simulation based in NS-3. They used a point to point NS-3 module to create a
+network of 2 levels of orbiting satellites and fixed points on the earth which
+acts as uplinks to the network. The paper analyzed TCP HighSpeed against TCP
+Vegas slow start. However this paper stops at expanding the network beyond earth
+orbit, and rather focuses on Earth orbit communication.
 
@@
 distributed across the surface or in the orbits of different planets or other
-celestial objects. In the year of 2024 the current interplanetary is limited to
-Mars, Earth, Earth orbit, Moon orbit, and satellites orbiting the sun. The
-interplanetary network is also expanded to deep space satellites, like the
-Voyager probes. A sample message path across the network could be that a message
-is passed from a large network antenna on earth to a earth orbiting satellite,
-then when it has direct line of sight with MAVEN (a satellite orbiting mars) it
-sends the message. MAVEN then relays the message to a Mars rover like Odyssey.
-This description hides the complexity which arises in an interplanetary network
-which may seem like a extremely simple network.
+celestial objects. In the year of 2024 the current interplanetary network is
+limited to Mars, Earth, Earth orbit, Moon orbit, and satellites orbiting the
+sun. The interplanetary network is also expanded to deep space satellites, like
+the Voyager probes. A sample message path across the network could be that a
+message is passed from a large network antenna on earth to a earth orbiting
+satellite, then when it has direct line of sight with MAVEN (a satellite
+orbiting mars) it sends the message. MAVEN then relays the message to a Mars
+rover like Odyssey. This description hides the complexity which arises in an
+interplanetary network which may seem like a extremely simple network.
 
-In a traditional network a majority of the network traffic will occur over a
+In a traditional network, a majority of the network traffic will occur over a
 wired connection, especially for transfers of large amounts of data. This means
-that the error is relatively low. This is no longer the case in the case of
-satellites sending messages over potentially millions of kilometers. Atmosphere
-in the Earth orbit, solar radiation, and dust cause increased error in sending
-the message between entities. Inverse square law dictates that over these large
+that the error is relatively low. This is no longer the case with satellites
+sending messages over potentially millions of kilometers. Atmosphere in the
+Earth orbit, solar radiation, and dust cause increased error in sending the
+message between entities. Inverse square law dictates that over these large
 periods the energy needed to send the message so that it can be read at its
@@
 
-Other than physical limitations a interplanetary network has a rapidly changing
-topology at anytime. For instance satellites which orbit planets can have their
-field of view blocked as they move over the horizon. While terrestrial networks
-also have rapidly changing networks at the edges, but the difference is that the
-core of the network is often stable. Where depending on the hour, day, year, the
-core of the interplanetary network can change drastically. To be able to connect
-to the network there are no well known places to access the network, except for
-Earth.
+Other than physical limitations an interplanetary network has a rapidly changing
+topology at all times. For instance, satellites which orbit planets can have
+their field of view blocked as they move over the horizon. While terrestrial
+networks also have rapidly changing networks at the edges, the core of the
+network is often stable. Where depending on the hour, day, year, the core of the
+interplanetary network can change drastically. To be able to connect to the
+network there are no well known places to access the network, except for Earth.
 
@@
 confirmations. Other protocols like store and load, and Licklider are used to
-also overcome these problems. In this paper since no complete implementation of
-SCPS TP has been created for NS-3, New Reno TCP was used instead.
+also overcome these problems. New Reno TCP was used instead of SCPS TP since no
+complete implementation of SCPS TP has been created for NS-3.
 
@@
 orbit around the sun. The orbital radius was chosen by taking the average
-orbital radius of the real life celestial bodies. For example Earth's average
-orbital radius is 149.6 billion meters~\cite{nasa_earth}, and therefore orbits
-around the sun in a circular orbit of radius 149.6 billion meters in our model.
-Another result of this assumption is that orbits are considered to be centered
-around the body which it orbits. The Mars orbit has a focal point that is
-further in space, making it more oval shaped than the earth
+orbital radius of the real life celestial bodies. For example, Earth's average
+orbital radius is 149.6 billion meters~\cite{nasa_earth} and therefore, it
+orbits around the sun in a circular orbit of radius 149.6 billion meters in our
+model. Another result of this assumption is that orbits are considered to be
+centered around the body which it orbits. The Mars orbit has a focal point that
+is further in space, making it more oval shaped than the earth
 orbit.~\cite{nasa_mars} While this could make changes to a simulated network in
 some very specific situations, the primary behaviour of a solar system is
-maintained, like retrograde orbits, periodical alignment, predictable behaviour,
-and recursive orbits of sub orbiting object.
+maintained such as retrograde orbits, periodical alignment, predictable
+behaviour, and recursive orbits of sub orbiting object.
 
@@
 
-Based on these simplifications, a physical simulation with several goals. To
-calculate if a line of sight could be made between two communicating entities,
-that is if a message can be sent, to calculate the error rate when sending a
-message, and to calculate the message propagation delay for the message to be
-sent. Line of sight was achieved using simple geometry calculations. All
-entities are able to block signals if the ray passes through the radius of the
-entity. One of the largest factors contributing to transmission error is the
-distance traveled by the electromagnetic signal between transmitting and
-receiving nodes. By generalizing Friis' transmission formula\cite{Friis} for
-isotropic emitters and receivers, we can obtain the ratio of signal intensity
-received to signal intensity transmitted, known as Free Space Path Loss (FSPL).
-We can use FSPL as our transmission error rate, giving the following expression:
-$Error Rate = {(\frac{4 \pi d}{\lambda})}^2$ where $d$ is the distance between
-nodes in metres and $\lambda$ is the wavelength of the electromagnetic wave used
-in transmission. As much of space communications utilizes the Ka-band of
-frequencies (27--40 GHz)\cite{Morabito_Hastrup} we opted to use a transmission
-frequency of 30Ghz. The transmission wavelength can be then calculated according
-to $\lambda = \frac{c}{f}$, where $c$ is the speed of light and $f$ is the
+We designed a physical simulation with several goals based on these
+simplifications. To calculate if a line of sight could be made between two
+communicating entities, that is if a message can be sent, to calculate the error
+rate when sending a message, and to calculate the message propagation delay for
+the message to be sent. Line of sight was achieved using simple geometry
+calculations. All entities are able to block signals if the ray passes through
+the radius of the entity. One of the largest factors contributing to
+transmission error is the distance traveled by the electromagnetic signal
+between transmitting and receiving nodes. By generalizing Friis' transmission
+formula~\cite{Friis} for isotropic emitters and receivers, we can obtain the
+ratio of signal intensity received to signal intensity transmitted, known as
+Free Space Path Loss (FSPL). We can use FSPL as our transmission error rate,
+giving the following expression: $Error Rate = {(\frac{4 \pi d}{\lambda})}^2$
+where $d$ is the distance between nodes in metres and $\lambda$ is the
+wavelength of the electromagnetic wave used in transmission. As much of space
+communications utilizes the Ka-band of frequencies (27--40
+GHz)~\cite{Morabito_Hastrup}, we opted to use a transmission frequency of 30Ghz.
+The transmission wavelength can be then calculated according to
+$\lambda = \frac{c}{f}$, where $c$ is the speed of light and $f$ is the
 transmission frequency, giving a transmission wavelength of about 9.99~mm.
@@
 Station, Mars, and the Mars Orbiter. All were able to connect to the network
-except for the the Sun and the Moon. The input of the physics engine is a time
-$t$ from some preset initial conditions which outputs the state of all entities.
+except for the Sun and the Moon. The input of the physics engine is a time $t$
+from some preset initial conditions which outputs the state of all entities.
 Which is fed into the NS-3 simulation.
@@
 NS-3 is a useful library, but we did run into a number of problems over the
-course of our implementation of our simulation. The first problem came when we
+course of our implementation of the simulation. The first problem came when we
 first tried to compile the library itself and ran into a bug in the build script
@@
 
-During the work on this project we contributed to the open source NS-3 project 
-by committing a change to fix a bug which occurred when running the program on 
-different operating systems.
-
 \section{Data}
@@
 
-Data was scraped from the trace output that was generated while running the 
-ns3 simulation. Trace data was given when a message was queued, dequeued, 
-and received at the destination. This data also included the amount of 
-data being sent in bytes. There is other data on the messages being sent, but 
-was not tracked for the purpose of this project. The trace data was parsed 
-using regex expressions to extract the needed numbers from the string.
-This data was then fed into a pandas dataframe. A unique id had to be 
-generated for each message since once a message was received the id was 
-reused. This was done by tracking when each message was received with id 
-$x$ and after incrementing a number by one, say $y$. The unique id was then 
-created by hashing $xy$. Once the data frame was created it had to be merged 
-on itself to have one row of the dataframe represent exactly one message rather 
-then one row be either queue, dequeue, or received.
+Data was scraped from the trace output that was generated while running the NS-3
+simulation. Trace data was given when a message was queued, dequeued, or
+received at the destination. This data also included the amount of data being
+sent in bytes. There is other data on the messages being sent, but was not
+tracked for the purpose of this project. The trace data was parsed using regular
+expressions to extract the needed numbers from the string. This data was then
+fed into a \texttt{pandas} dataframe. A unique ID had to be generated for each
+message since once a message was received the ID was reused. This was done by
+tracking when each message was received with ID $x$ and after incrementing a
+number by one, say $y$. The unique ID was then created by hashing $xy$. Once the
+data frame was created it had to be merged on itself to have one row of the
+dataframe represent exactly one message rather then one row be either queue,
+dequeue, or received.
 
@@
 
-The message size is the same across all of the messages. This makes sense as the message size do not need to be different across different messages. This is different compared to TCP and New Reno.
+The message size is the same across all of the messages. This makes sense as the
+message size do not need to be different across different messages. This is
+different compared to TCP and New Reno.
 
-On the left side of the graph before approximately $t = 25000$ this is where
-the planet most likely has a direct path between Earth and Mars. This has a
-steady pattern of a batch of messages are queued, they are sent, and after they
-are sent the next batch of messages are queued, and calculated the propagation
-delay at the time that it is queued. This leads to the clusters which can be seen.
-The curvature of the delay is most likely attributed to the movement of Earth
-relative to Mars, that is that they are getting closer together. Queueing delay is likely small in this section suggested by the tight groupings.
+On the left side of the graph, before about $t = 25000$ seconds is where there
+most likely is a direct path between Earth and Mars. This has a steady pattern
+of a batch of messages are queued, they are sent, and after they are sent the
+next batch of messages are queued, and calculated the propagation delay at the
+time that it is queued. This leads to the clusters which can be seen. The
+curvature of the delay is most likely attributed to the movement of Earth
+relative to Mars, that is that they are getting closer together. Queuing delay
+is likely small in this section suggested by the tight groupings.
 
-At approximately time $t = 25000$ the direct path between Earth and Mars is
+At approximately $t = 25000$ seconds, the direct path between Earth and Mars is
 blocked. This means that the messages have to be routed through a secondary
@@
 
-This analysis cannot be 100\% verified due to the data mined from NS-3 did not give a ton of information about the messages which were sent on our simulated network.
+This analysis cannot be 100\% verified due to the data mined from NS-3 did not
+give as much information as we would have liked about the messages which were
+sent on our simulated network.
 
@@
 difference in the amounts of bytes transferred in an ACK message compared to
-when data is being sent.
+when data is being sent. On average, the queue delay was negligible. The total
+transfer time was 261.56 seconds on average.
 
-Unlike in UDP the motion of the planets is not observable in the graph. However
+Unlike in UDP, the motion of the planets is not observable in the graph. However
 this is because there are several different paths which the TCP message could
@@
 congestion~\cite{Nahar2016}, was used in place of SCP TP. New Reno behaved very
-similar to TCP. Its average data transfer was 307.18 bytes.
-
-It preformed marginally better in its time from being queued to the message
-received. This was mostly due to a decrease in time that it was in queue, which
-is to be expected. There was a very marginal increase in time in transit
-however.
+similar to TCP. Its average data transfer was 307.18 bytes. On average, the
+queue delay was $8.69\times10^{-4}$ seconds. The average total transfer time was
+261.56 seconds. The differences in our simulation between this and TCP were
+negligible.
 
@@
 
-Similar to TCP we can see the two stages which occur in the simulation. One
-where most of the messages were likely being routed from directly from Earth to
+Similar to TCP we can see the two stages which occur in the simulation. One,
+where most of the messages were likely being routed directly from Earth to
 Mars, which then gets transformed into having to route through one of the
@@
 
-Since a large amount of propagation delay dominates the amount of time that it
-takes for messages total transmit time from being queued to being received at
-the appropriate node, other increases in performance mean only a slight decrease
-in time sending the message. An important statistic to pay attention to is that
-UDP almost always sent a message with a size of 1052 bytes. This means that even
-though TCP, New Reno, and UDP had sent a similar number of messages, the actual
-throughput for UDP was significantly higher, which is much more significant in
-an IPN situation. In the data graphs we can see that not only was the maximum
-data being sent was less than UDP, but also had to send messages of very little
-size, most likely acknowledgments, as well. New Reno was seen to improve on the
-queuing time compared to both TCP and UDP. Which would mean in a high
-reliability situation and in situations where higher traffic is sent through a
-network with a relatively low bandwidth, like a IPN, where satellites are
-limited by how much data they can transmit, New Reno's effect can be more
-significant. This should only be used in a connection where propagation delay is
-less of an issue like between planet and satellite communication. Queuing time
-in UDP took a more significant percentage of its time than both TCP and New
-Reno. In general however, UDP is the best option for primary communication
-between Mars and Earth communication.
+Since a large amount of propagation delay dominates the total transmit time from
+being queued, to being received at the appropriate node, other increases in
+performance mean only a slight decrease in time sending the message. An
+important statistic to pay attention to is that UDP almost always sent a message
+with a size of 1052 bytes. This means that even though TCP, New Reno, and UDP
+had sent a similar number of messages, the actual throughput for UDP was
+significantly higher, which is much more significant in an IPN situation. In the
+data graphs we can see that not only was the maximum data being sent was less
+than UDP, but also had to send messages of very little size, most likely
+acknowledgments, as well. Queuing time in UDP took a more significant percentage
+of its time than both TCP and New Reno. In general however, UDP is the best
+option for primary communication between Mars and Earth communication.
 
@@
 TP as a network protocol which could be used and compared to the protocols
-analyzed in this paper. Other protocols like the Linklider or load and store
-protocols could also be implemented, to create a more realistic IPN.
+analyzed in this paper. The Linklider or store and load protocols could also be
+implemented, to create a more realistic IPN.
 
@@
 
-In this paper we outlined how we created a model of an interplanetary internet
-(IPN) using a physics engine which fed into a NS-3 simulation. The physics
-simulation used a simplified model of orbiting planets and satellites as well as
-physics calculations to create data which would be the output for any input time
-$t$. The NS-3 model used a point to point base which programmatically took the
-data from the physics engine and created nodes which acted as routers. A
-simulation was ran which transferred data between Earth and Mars, using
-satellites and other entities to relay messages around blocked connections.
-Analysis was done on three different network protocols used in the IPN, that is
-UDP, TCP and New Reno. UDP was found to have the most throughput out of all of
-the three protocols, while New Reno reduced the amount of time a message was
-queued.
+In this paper, we outlined how we created a model of an interplanetary internet
+using a physics engine which fed into a NS-3 simulation. The physics simulation
+used a simplified model of orbiting planets and satellites as well as physics
+calculations to create data which would be the output for any input time $t$.
+The NS-3 model used a point to point base which programmatically took the data
+from the physics engine and created nodes which acted as routers. A simulation
+was ran which transferred data between Earth and Mars, using satellites and
+other entities to relay messages around blocked connections. Analysis was done
+on three different network protocols used in the IPN, that is UDP, TCP and New
+Reno. UDP was found to have the most throughput out of all of the three
+protocols.
 

### 2024-04-10 process_data.py

@@ -184,5 +184,5 @@ def save_dataframe():
 
 # for testing
 save_dataframe()
-get_graphs()
+# get_graphs()
 get_statistics()
